---
title: "main-analysis"
author: "YG"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.width = 12,
                      fig.height = 6,
                      dev = "png",
                      message = FALSE, 
                      warning = FALSE, 
                      error=TRUE, 
                      cache = FALSE,
                      include = TRUE)

# set a CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))
```


```{r libraries, echo=FALSE, include=FALSE}
# clean all cached data
rm(list = ls()) 

# for getting standard beta hat and CI
library(insight)

library(report)

# load all relevant packages for data tidying, analysis and plotting (e.g., dplyr, lubridate -- to convert date and time data)
library(tidyverse) 

# provide a mechanism for chaining commands with a new forward-pipe operator, %>%
library(magrittr)

# dependency package for lme4 
library(Matrix)

# fit linear and generalized linear mixed-effects models
library(lme4)

# provide p-values in type I, II or III anova and summary tables for lmer model fits (cf. lme4) via Satterthwaite's degrees of freedom method
library(lmerTest)

# flexibly restructure and aggregate data
library(reshape2)

#library(logistf)

# provide some easy-to-use functions for creating and customizing 'ggplot2'- based publication ready plots
# but we could set this aside at this point 
# library(ggpubr)

# provide geoms for ggplot2 to repel overlapping text labels
# for latest features, download and install directly from git 
# devtools::install_github("slowkow/ggrepel")
library(ggrepel)

# provide a number of user-level functions to work with "grid" graphics, notably to arrange multiple grid-based plots on a page, and draw tables
library(gridExtra)
library(cowplot)
# reorder factor levels (including moving specified levels to front, ordering by first appearance, reversing, and randomly shuffling); modifies factor levels (including collapsing rare levels into other, 'anonymising', and manually 'recoding')
library(forcats)

# create interactive web graphics from 'ggplot2' graphs 
library(plotly)

# create US map 
library(usmap)

# generate summary table for the mixed effects model 
library(sjPlot)

# transform the data with matching strings 
library(stringr)

# check for multicolinearity among predictors
library(car)

# detect changes in the models and avoids redundant rerunning of models 
library(digest)

# beeswarm plots
library(ggbeeswarm)

library(Hmisc)
```


```{r constants, echo=FALSE, include=FALSE}
# aesthetic defaults

dodge_amt <- 0.9
dodge_amt2 <- 0.5

# set the theme for graphs 
theme_set(
  theme_minimal(base_size = 20) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold")))
  # panel.background = element_rect(fill = "transparent")))
```


### 1. Transcription data analysis
```{r scoring-functions, echo=FALSE}
##### helper function
# clean special characters
chars <- c("<", ">", '"', ",", ";", "|", "'")
char_codes <- c("&lt&", "&gt&", "&quot&", "&comma&", "&semicol&", "&pipe&", "&apos&")


mgsub <- function(pattern, replacement, x, ...) {
  if (length(pattern) != length(replacement)) {
    stop("pattern and replacement do not have the same length.")
  }
  result <- x
  for (i in 1:length(pattern)) {
    result <- gsub(pattern[i], replacement[i], result, ...)
  }
  result
}

# keyword scoring function
num.keywords.correct <- function(transcription, keywords) {
  keys <- gsub("[[:punct:]]", "", tolower(as.character(keywords)))
  trans <- gsub("[[:punct:]]", "", tolower(as.character(transcription)))
  keys.list <- unlist(strsplit(keys, " "))
  trans.list <- unlist(strsplit(trans, " "))
  return(sum(!is.na(pmatch(keys.list, trans.list))))
}

# keyword checking (for Plot 6 and 7)
check_keywords_correct <- function(transcription, keywords, trial_id, filename, participant_id) {
  keys <- gsub("[[:punct:]]", "", tolower(as.character(keywords)))
  trans <- gsub("[[:punct:]]", "", tolower(as.character(transcription)))
  keys.list <- unlist(strsplit(keys, " "))
  trans.list <- unlist(strsplit(trans, " "))
  correct_matches <- keys.list %in% trans.list
  return(data.frame(
    trial_id = trial_id,
    participant_id = participant_id,
    Filename = filename,
    keyword = keys.list,
    is_correct = correct_matches
  ))
}


##### main processing function

process_transcription_data <- function(data_filepath, keyword_filepath) {
  
  d <- read.csv(data_filepath)
  key.df <- read.delim(keyword_filepath, header = TRUE, sep = ",")
  key.df$Sentence <- as.character(key.df$Sentence)
  
  d.trans <- subset(d, !is.na(stimulus_files_presented) & stimulus_files_presented != "" & block == "Transcription")
  
  d.trans$Filename <- basename(as.character(d.trans$stimulus_files_presented))
  d.trans$Transcription <- d.trans$response_value
  
  cols_to_keep <- c("participant_id", "Filename", "Transcription", "block", "trial_num_sequential")
  
  d.trans_subset <- d.trans[, names(d.trans) %in% cols_to_keep]

  d.all.m <- merge(d.trans_subset, subset(key.df, select = -c(SentenceSet)),
                   by.x = "Filename", by.y = "Sentence",
                   sort = FALSE)

  d.all.m <- within(d.all.m, {
    Transcription <- tolower(mgsub(char_codes, chars, Transcription))
    NumKeywordsCorrect <- mapply(num.keywords.correct, Transcription, Keywords)
    PropKeywordsCorrect <- NumKeywordsCorrect / NumKeywords
  })
  
  print(paste("--- Scoring complete for:", basename(data_filepath), "---"))
  print(summary(d.all.m$PropKeywordsCorrect))
  
  return(d.all.m)
}


##### sanity check function
perform_sanity_checks <- function(scored_data) {
  
  # 1. Tally N of unique subjects
  unique_subject_count <- scored_data %>%
    summarise(N_Unique_Subjects = n_distinct(participant_id))
  
  print("--- Sanity Check 1: Total Unique Subjects ---")
  print(unique_subject_count)
  cat("\n")
  
  # 2. Tally N of unique sentences per subject
  sentences_per_subject <- scored_data %>%
    group_by(participant_id) %>%
    summarise(N_Unique_Sentences = n_distinct(Filename))
  
  print("--- Sanity Check 2: Unique Sentences Transcribed Per Subject ---")
  print(as.data.frame(sentences_per_subject))
  cat("\n")
}


##### plot generation function
generate_analysis_plots <- function(scored_data) {

  
  # overall summary stats
  summary_stats <- scored_data %>%
    summarise(
      Mean = mean(PropKeywordsCorrect),
      Median = median(PropKeywordsCorrect),
      SD = sd(PropKeywordsCorrect),
      Min = min(PropKeywordsCorrect),
      Max = max(PropKeywordsCorrect)
    )
  
  # overall distribution
  mean_score <- mean(scored_data$PropKeywordsCorrect)
  p1 <- ggplot(scored_data, aes(x = PropKeywordsCorrect)) +
    geom_histogram(binwidth = 0.1, fill = "steelblue", color = "white", alpha = 0.9) +
    geom_vline(xintercept = mean_score, color = "darkorange", linetype = "dashed", size = 1.2) +
    annotate("text", x = mean_score - 0.02, y = 14, label = paste("Mean =", round(mean_score, 2)),
             color = "darkorange", hjust = "right", fontface = "bold", size = 4) +
    labs(title = "Distribution of Overall Transcription Performance",
         x = "Proportion Keywords Correct", y = "Number of Trials") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  # subj-level 
plot_data <- scored_data %>%
    group_by(participant_id, block) %>%
    summarise(Mean_Prop_Correct = mean(PropKeywordsCorrect), .groups = "drop")


summary_data <- plot_data %>%
    group_by(block) %>%
    summarise(stats = list(smean.cl.boot(Mean_Prop_Correct)), .groups = "drop") %>%
    tidyr::unnest_wider(stats)


p2 <- ggplot(summary_data, aes(x = block, y = Mean)) +
    geom_point(color = "darkred", size = 4) +
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "darkred", linewidth = 1, width = 0.2) +
    labs(title = "Transcription Performance (Mean + 95% CI)",
         x = "Block", y = "Proportion Keywords Correct") +
    theme_bw() +
    theme(axis.ticks.x = element_line(), plot.title = element_text(size = 16, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))

  
# item level
  item_means <- scored_data %>%
    group_by(Filename, NumKeywords) %>%
    summarise(Mean_Prop_Correct = mean(PropKeywordsCorrect), .groups = "drop") %>%
    ungroup() %>%
    arrange(Mean_Prop_Correct)
  
  mean_item_difficulty <- mean(item_means$Mean_Prop_Correct)
  
  p3 <- ggplot(item_means, aes(x = Mean_Prop_Correct)) +
    geom_histogram(binwidth = 0.1, fill = "mediumpurple", color = "white", alpha = 0.9) +
    geom_vline(xintercept = mean_item_difficulty, color = "gold", linetype = "dashed", size = 1.2) +
    annotate("text", x = mean_item_difficulty + 0.02, y = 6.5, 
             label = paste("Mean =", round(mean_item_difficulty, 2)),
             color = "gold", hjust = "left", fontface = "bold", size = 6) +
    labs(title = "Distribution of Item Difficulty",
         x = "Mean Proportion Correct (per Sentence)", y = "Number of Sentences") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  item_means_clean <- item_means %>%
    mutate(Filename_clean = gsub(".wav", "", Filename))
  
  p4 <- ggplot(item_means_clean, aes(x = Mean_Prop_Correct, y = reorder(Filename_clean, Mean_Prop_Correct))) +
    geom_segment(aes(x = 0, xend = Mean_Prop_Correct,
                     y = reorder(Filename_clean, Mean_Prop_Correct), 
                     yend = reorder(Filename_clean, Mean_Prop_Correct)),
                 color = "grey") +
    geom_point(aes(color = Mean_Prop_Correct), size = 3) +
    scale_color_gradient(low = "red", high = "green", name = "Mean Score") +
    labs(title = "Difficulty of Individual Sentences",
         x = "Mean Proportion Correct", y = "Sentence Filename") +
    theme_bw() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  # confound check: are longer sentences more difficult?
  p5 <- ggplot(scored_data, aes(x = NumKeywords, y = PropKeywordsCorrect)) +
    geom_point(alpha = 0.2) +
    geom_jitter(width = 0.2, alpha = 0.3) +
    geom_smooth(method = "lm", color = "red") + 
    labs(title = "Performance by Number of Keywords",
         x = "Number of Keywords in Sentence", y = "Proportion Keywords Correct") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  # keyword in sentence accuracy
  scored_data$trial_id <- 1:nrow(scored_data)
  keyword_long_df <- pmap_dfr(
    list(
      scored_data$Transcription,
      scored_data$Keywords,
      scored_data$trial_id,
      scored_data$Filename,
      scored_data$participant_id
    ),
    check_keywords_correct 
  )
  
  keyword_accuracy <- keyword_long_df %>%
    group_by(Filename, keyword) %>%
    summarise(Prop_Correct = mean(is_correct), .groups = "drop") %>%
    ungroup()
  
  p6 <- ggplot(keyword_accuracy, aes(x = reorder(keyword, Prop_Correct), 
                                   y = Prop_Correct, fill = Prop_Correct)) +
    geom_col(color = "black") + 
    facet_wrap(~ Filename, scales = "free_x") + 
    scale_fill_gradient(low = "mediumpurple", high = "gold", name = "Accuracy") +
    labs(title = "Accuracy for Individual Keywords in Each Sentence",
         x = "Keyword (Sorted by Accuracy)",
         y = "Proportion Correct (across all participants)") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 12),
          plot.title = element_text(size = 16, face = "bold"),
          strip.text = element_text(size = 12, face = "bold"))
  
  # hardest keywords
  overall_keyword_accuracy <- keyword_long_df %>%
    group_by(keyword) %>%
    summarise(
      N_Occurrences = n(),
      Prop_Correct = mean(is_correct)
    ) %>%
    ungroup() %>%
    arrange(Prop_Correct)
  
  p7 <- ggplot(head(overall_keyword_accuracy, 20), aes(x = reorder(keyword, Prop_Correct), y = Prop_Correct)) +
    geom_col(fill = "mediumpurple", color = "white", alpha = 0.7) +
    labs(title = "Top 20 Hardest Keywords (Across All Sentences)",
         x = "Keyword", y = "Overall Proportion Correct") +
    coord_flip() + 
    theme_bw() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  plot_list <- list(p1 = p1, p2 = p2, p3 = p3, p4 = p4, p5 = p5, p6 = p6, p7 = p7)
  
  return(plot_list)
}


# accuracy over time function
generate_time_analysis_plots <- function(scored_data) {
  
 d.with_trial_order <- scored_data %>%
    arrange(participant_id, trial_num_sequential) %>%
    rename(trial_order = trial_num_sequential)
  
  # accuracy over time 
  transcription_summary_over_time <- d.with_trial_order %>%
    group_by(trial_order) %>%
    summarise(
      stats = list(smean.cl.boot(PropKeywordsCorrect)),
      .groups = "drop"
    ) %>%
    tidyr::unnest_wider(stats)
  
  p8 <- ggplot(transcription_summary_over_time, aes(x = trial_order, y = Mean)) +
    geom_point(color = "darkblue", size = 3) +
    geom_line(color = "darkblue", linetype = "dashed", alpha = 0.5) +
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "darkblue", width = 0.2, linewidth = 0.8) +
    geom_smooth(
      data = d.with_trial_order, 
      aes(x = trial_order, y = PropKeywordsCorrect), 
      method = "lm", se = TRUE, color = "firebrick", fill = "firebrick",
      alpha = 0.1, inherit.aes = FALSE
    ) +
    scale_x_continuous(breaks = seq(0, 20, by = 2)) + # **CORRECTED SYNTAX**
    scale_y_continuous(limits = c(0, 1.0)) + 
    labs(
      title = "Transcription Accuracy Over Time (Group Average)",
      x = "Trial Number",
      y = "Mean Proportion Keywords Correct"
    ) +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.title = element_text(size = 12), axis.text = element_text(size = 12))
  
  # individual accuracy over time
  p9 <- ggplot(d.with_trial_order, aes(x = trial_order, y = PropKeywordsCorrect)) +
    geom_point(alpha = 0.4, color = "steelblue") +
    geom_smooth(method = "lm", se = FALSE, color = "firebrick", linewidth = 1) +
    facet_wrap(~ participant_id, ncol = 4) +
    scale_x_continuous(breaks = c(1, 10, 20)) + 
    scale_y_continuous(limits = c(0, 1.0)) + 
    labs(
      title = "Individual Participant Transcription Accuracy Over Time",
      x = "Trial Number",
      y = "Proportion Keywords Correct"
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      axis.title = element_text(size = 12),
      strip.text = element_text(size = 10, face = "bold", color = "white"),
      strip.background = element_rect(fill = "steelblue")
    )

  plot_list <- list(p8 = p8, p9 = p9)
  return(plot_list)
}
```


#### 1.1 Baseline transcription data analysis
```{r transcription-data-analysis-baseline}
d.transcrip.baseline.all <- process_transcription_data(
  data_filepath = "../data/preprocessed/transcrip_baseline/transcription_data.csv",
  keyword_filepath = "../data/transcription_scoring.txt"
)

# sanity checks
perform_sanity_checks(d.transcrip.baseline.all)

# plots
all_plots_transcrip_baseline <- generate_analysis_plots(d.transcrip.baseline.all)


time_plots_transcrip_baseline <- generate_time_analysis_plots(d.transcrip.baseline.all)
```

```{r p1-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 1: Overall Distribution of Transcription Performance - Baseline"}

print(all_plots_transcrip_baseline$p1)
```

```{r p2-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 2: Transcription Performance by Participant - Baseline"}
print(all_plots_transcrip_baseline$p2)
```

```{r p3-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 3: Distribution of Item Difficulty - Baseline"}
print(all_plots_transcrip_baseline$p3)
```


```{r p4-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 4: Difficulty of Individual Sentences - Baseline"}
print(all_plots_transcrip_baseline$p4)
```


```{r p5-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 5: Performance by Number of Keywords - Baseline"}
print(all_plots_transcrip_baseline$p5)
```


```{r p6-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 6: Accuracy for Individual Keywords in Each Sentence - Baseline"}
print(all_plots_transcrip_baseline$p6)
```


```{r p7-analysis-plots-transcription-baseline, echo=FALSE, fig.cap="Figure 7: Top 20 Hardest Keywords (Across All Sentences) - Baseline"}
print(all_plots_transcrip_baseline$p7)
```


```{r p8-analysis-plot-transcription-baseline, echo=FALSE, fig.cap="Figure 8: Transcription Performance by Trial Order - Baseline"}
print(time_plots_transcrip_baseline$p8)
```

```{r p9-analysis-plot-transcription-baseline, echo=FALSE, fig.cap="Figure 9: Individual Participant Performance by Trial Order - Baseline"}
print(time_plots_transcrip_baseline$p9)
```


#### 1.2 Main group transcription data analysis

##### 1.2.1 Main batch 1 - 48 subjects (first pass only)

```{r transcription-data-analysis-main}
d.transcrip.main.batch1.48subj <- process_transcription_data(
  data_filepath = "../data/preprocessed/main_batch1/48subj_transcription_data.csv",
  keyword_filepath = "../data/transcription_scoring.txt"
)

# sanity checks
perform_sanity_checks(d.transcrip.main.batch1.48subj)

# plots
all_plots_transcrip_main_batch1_48subj <- generate_analysis_plots(d.transcrip.main.batch1.48subj)


time_plots_transcrip_main_batch1_48subj <- generate_time_analysis_plots(d.transcrip.main.batch1.48subj)
```

```{r p1-analysis-plots, echo=FALSE, fig.cap="Figure 1: Overall Distribution of Transcription Performance - Main-Batch1-48subj"}

print(all_plots_transcrip_main_batch1_48subj$p1)
```

```{r p2-analysis-plots, echo=FALSE, fig.cap="Figure 2: Transcription Performance by Participant - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p2)
```

```{r p3-analysis-plots, echo=FALSE, fig.cap="Figure 3: Distribution of Item Difficulty - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p3)
```


```{r p4-analysis-plots, echo=FALSE, fig.cap="Figure 4: Difficulty of Individual Sentences - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p4)
```


```{r p5-analysis-plots, echo=FALSE, fig.cap="Figure 5: Performance by Number of Keywords - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p5)
```


```{r p6-analysis-plots, echo=FALSE, fig.cap="Figure 6: Accuracy for Individual Keywords in Each Sentence - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p6)
```


```{r p7-analysis-plots, echo=FALSE, fig.cap="Figure 7: Top 20 Hardest Keywords (Across All Sentences) - Main-Batch1-48subj"}
print(all_plots_transcrip_main_batch1_48subj$p7)
```


```{r p8-analysis-plot, echo=FALSE, fig.cap="Figure 8: Transcription Performance by Trial Order - Main-Batch1-48subj"}
print(time_plots_transcrip_main_batch1_48subj$p8)
```

```{r p9-analysis-plot, echo=FALSE, fig.cap="Figure 9: Individual Participant Performance by Trial Order - Main-Batch1-48subj"}
print(time_plots_transcrip_main_batch1_48subj$p9)
```


##### 1.2.2 Main batch 3 - 3 subjects (first and second pass)

```{r transcription-data-analysis-main}
d.transcrip.main.batch3.3subj <- process_transcription_data(
  data_filepath = "../data/preprocessed/main_batch3/testrun_4subj_transcription_data.csv",
  keyword_filepath = "../data/transcription_scoring.txt"
)

# we are excluding one subj data bc that subj did not engage with the main error detection task
d.transcrip.main.batch3.3subj <- d.transcrip.main.batch3.3subj %>%
  filter(participant_id != "6df14d8680851e21a1bbe20d8e32fc04bc30979c5d654dc35ef36bc2910b9765")

# sanity checks
perform_sanity_checks(d.transcrip.main.batch3.3subj)

# plots
all_plots_transcrip_main_batch3_3subj <- generate_analysis_plots(d.transcrip.main.batch3.3subj)


time_plots_transcrip_main_batch3_3subj <- generate_time_analysis_plots(d.transcrip.main.batch3.3subj)
```

```{r p1-analysis-plots, echo=FALSE, fig.cap="Figure 1: Overall Distribution of Transcription Performance - Main-Batch3-3subj"}

print(all_plots_transcrip_main_batch3_3subj$p1)
```

```{r p2-analysis-plots, echo=FALSE, fig.cap="Figure 2: Transcription Performance by Participant - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p2)
```

```{r p3-analysis-plots, echo=FALSE, fig.cap="Figure 3: Distribution of Item Difficulty - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p3)
```


```{r p4-analysis-plots, echo=FALSE, fig.cap="Figure 4: Difficulty of Individual Sentences - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p4)
```


```{r p5-analysis-plots, echo=FALSE, fig.cap="Figure 5: Performance by Number of Keywords - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p5)
```


```{r p6-analysis-plots, echo=FALSE, fig.cap="Figure 6: Accuracy for Individual Keywords in Each Sentence - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p6)
```


```{r p7-analysis-plots, echo=FALSE, fig.cap="Figure 7: Top 20 Hardest Keywords (Across All Sentences) - Main-Batch3-3subj"}
print(all_plots_transcrip_main_batch3_3subj$p7)
```


```{r p8-analysis-plot, echo=FALSE, fig.cap="Figure 8: Transcription Performance by Trial Order - Main-Batch3-3subj"}
print(time_plots_transcrip_main_batch3_3subj$p8)
```

```{r p9-analysis-plot, echo=FALSE, fig.cap="Figure 9: Individual Participant Performance by Trial Order - Main-Batch3-3subj"}
print(time_plots_transcrip_main_batch3_3subj$p9)
```

#### 1.3 Comparative analysis between baseline and main group


```{r compare-baseline-with-main}
d.transcrip.baseline.all$group <- "baseline"
d.transcrip.main.batch1.48subj$group <- "main"

common_cols <- intersect(names(d.transcrip.baseline.all), names(d.transcrip.main.batch1.48subj))
d.combined.transcrip <- rbind(
  d.transcrip.baseline.all[, common_cols], 
  d.transcrip.main.batch1.48subj[, common_cols]
)


d.combined.transcrip$group <- as.factor(d.combined.transcrip$group)


print(
  d.combined.transcrip %>%
    group_by(group) %>%
    summarise(
      N_Participants = n_distinct(participant_id),
      Mean_Accuracy = mean(PropKeywordsCorrect),
      SD_Accuracy = sd(PropKeywordsCorrect),
      N_Trials = n()
    )
)


summary_combined_transcrip <- d.combined.transcrip %>%
  group_by(group) %>%
  summarise(
    stats = list(smean.cl.boot(PropKeywordsCorrect)),
    .groups = "drop"
  ) %>%
  tidyr::unnest_wider(stats)

print(summary_combined_transcrip)


p10_bootstrapped_ci <- ggplot(summary_combined_transcrip, aes(x = group, y = Mean, color = group)) +
  
  geom_jitter(
    data = d.combined.transcrip, 
    aes(x = group, y = PropKeywordsCorrect),
    width = 0.2, 
    alpha = 0.15, 
    height = 0,
    inherit.aes = FALSE 
  ) +
  
 
  geom_point(size = 4) +
  
 
  geom_errorbar(
    aes(ymin = Lower, ymax = Upper),
    width = 0.1,
    linewidth = 1.2
  ) +
  
  scale_y_continuous(limits = c(0, 1.0)) + 
  scale_x_discrete(labels = c("baseline" = "baseline", 
                              "main" = "main")) +
  labs(
    x = "Experimental Group",
    y = "Proportion Keywords Correct"
  ) +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 14, face = "bold"))

print(p10_bootstrapped_ci)


# group difference model
# m_group_comparison_transcrip <- lmer(PropKeywordsCorrect ~ group + (1 | participant_id), data = d.combined.transcrip)
# print(summary(m_group_comparison_transcrip))
```



```{r compare-accuracy-change-over-trials, echo=FALSE, message=FALSE, warning=FALSE}
d.combined.time.transcrip <- d.combined.transcrip %>%
  group_by(participant_id) %>%
  mutate(trial_order = row_number()) %>%
  ungroup()
  

time_summary_combined_transcrip <- d.combined.time.transcrip %>%
  group_by(group, trial_order) %>%
  summarise(
    stats = list(smean.cl.boot(PropKeywordsCorrect)),
    .groups = "drop"
  ) %>%
  tidyr::unnest_wider(stats)
  
p11 <- ggplot(time_summary_combined_transcrip, aes(x = trial_order, y = Mean, color = group, fill = group)) +
  geom_line(aes(linetype = group), linewidth = 1) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper), alpha = 0.2, linetype = 0) +
  geom_smooth(
    data = d.combined.time.transcrip, 
    aes(x = trial_order, y = PropKeywordsCorrect),
    method = "lm", 
    se = FALSE, 
    linewidth = 1.5
  ) +
  scale_x_continuous(breaks = seq(0, 20, by = 2)) +
  scale_y_continuous(limits = c(0, 1.0)) +
  labs(
    x = "Trial Number",
    y = "Mean Proportion Keywords Correct",
    color = "Group",
    fill = "Group",
    linetype = "Group"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12))

print(p11)

# m_accuracy_improvement_group <- lmer(
#   PropKeywordsCorrect ~ trial_order * group + (1 + trial_order | participant_id), 
#   data = d.combined.time.transcrip
# )
# print(summary(m_accuracy_improvement_group))
```


#### 1.4 Accent familiarity X Transcription accuracy between main group and baseline group
```{r accent-familiarity-transcription-accuracy, echo=FALSE, message=FALSE, warning=FALSE}

d_main_q <- read_csv("../data/preprocessed/main_batch1/48subj_questionnaire_data.csv")
d_baseline_q <- read_csv("../data/preprocessed/transcrip_baseline/questionnaire_data.csv")


d_main_q$group <- "main"
d_baseline_q$group <- "baseline"

d_long_combined_q <- bind_rows(d_main_q, d_baseline_q)


question_keys <- c(
  "exposure_spanish_accent",
  "exposure_spanish"
)

d_questionnaire_wide <- d_long_combined_q %>%
  filter(block == "Questionnaire") %>%
  filter(stimuli_presented %in% question_keys) %>%
  select(participant_id, group, stimuli_presented, response_value) %>%
  pivot_wider(
    names_from = stimuli_presented,
    values_from = response_value
  ) %>%
  distinct(participant_id, .keep_all = TRUE)


# get one mean transcription accuracy score per participant
# and aggregate over participants

d_acc_summary_transcrip <- d.combined.transcrip %>%
  group_by(participant_id, group) %>%
  summarise(
    Mean_Accuracy = mean(PropKeywordsCorrect, na.rm = TRUE),
    .groups = "drop"
  )

# join accuracy with questionnaire data
d_analysis_q_acc <- d_acc_summary_transcrip %>%
  left_join(d_questionnaire_wide, by = c("participant_id", "group"))


familiarity_levels <- c("Never", "Sometimes", "Half the time", "Often", "All the time")
d_analysis_q_acc <- d_analysis_q_acc %>%
  mutate(exposure_spanish_accent = factor(exposure_spanish_accent, 
                                          levels = familiarity_levels))

# group level
group_summary_fam_acc <- d_analysis_q_acc %>%
  filter(!is.na(exposure_spanish_accent)) %>% 
  group_by(group, exposure_spanish_accent) %>%
  summarise(
    N_Participants = n(),
    Mean_Accuracy = mean(Mean_Accuracy, na.rm = TRUE),
    SD_Accuracy = sd(Mean_Accuracy, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(group, exposure_spanish_accent)

print(group_summary_fam_acc)



# plot familiarity vs accuracy
p_fam_acc <- ggplot(d_analysis_q_acc, 
                     aes(x = exposure_spanish_accent, y = Mean_Accuracy)) +
  

  geom_jitter(aes(color = group), width = 0.2, alpha = 0.7, size = 2) +
  

  geom_boxplot(aes(fill = group), alpha = 0.2, outlier.shape = NA) +
  

  facet_wrap(~ group) +
  
 
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    x = "Self-Reported Familiarity with Spanish-Accented English",
    y = "Mean Keyword Accuracy"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 14, face = "bold")
  )


print(p_fam_acc)


```





### 2. Error detection data analysis

#### 2.1 Main batch1: 48 subjects data
```{r error-detection-annotated-data, message=FALSE, warning=FALSE}

# load data and exclude participants who failed the attention checks


tryCatch({
 
  df_fb <- read_csv("../data/annotated/main_batch1_first_pass_only_FB.csv")
  df_jb <- read_csv("../data/annotated/main_batch1_first_pass_only_JB.csv")
  df_quest <- read_csv("../data/preprocessed/main_batch1/48subj_questionnaire_data.csv")
  
  df_exp_all <- bind_rows(df_fb, df_jb)

  # attention checks (<6/8 OR 4/7)
  # one misannotated 4/7 in comments
  
  df_exp_all <- df_exp_all %>%
    mutate(
      
      comments = as.character(comments),
      
     
      attention_score_8 = as.numeric(
        str_extract(comments, "(?<=\\()\\d+(?=\\/8)")
      ),
      
     
      attention_score_7 = as.numeric(
        str_extract(comments, "(?<=\\()\\d+(?=\\/7)")
      )
    )


  failed_participants_8 <- df_exp_all %>%
    filter(attention_score_8 < 6) %>%
    pull(participant_id) %>%
    unique()
    
  failed_participants_7 <- df_exp_all %>%
    filter(attention_score_7 == 4) %>%
    pull(participant_id) %>%
    unique()
  
 
  failed_participants <- union(failed_participants_8, failed_participants_7)

  print(paste("Found", length(failed_participants), "total unique participants who failed."))
  print("Failed participant IDs:")
  print(failed_participants)


  df_exp_all <- df_exp_all %>%
    mutate(attention_check_passed = !participant_id %in% failed_participants)
  
 
# sanity checks after excluding those who failed the attention checks
  
  status_counts <- df_exp_all %>%
    distinct(participant_id, attention_check_passed) %>%
    count(attention_check_passed)
  
  print(status_counts)
 

# join combined annotated, cleaned error detection data with questionnaire data
  df_final_joined <- df_exp_all %>%
    left_join(df_quest, by = "participant_id")
})


df_clean_attchp <- df_final_joined %>%
  filter(attention_check_passed == TRUE) %>% 
  distinct(participant_id, question, .keep_all = TRUE)

print(paste("Analyzing", n_distinct(df_clean_attchp$participant_id), "participants who passed attention checks."))
```


```{r error-detection-data-plotting, message=FALSE, warning=FALSE}
# make sure the data is ready for plotting
# for example, examine whether there are any manual coding errors
df_analysis_err <- df_clean_attchp %>%
  # recode 'nanas' to 'nans'
  # bc of typo in manual coding
  mutate(
    err_reported_clean = case_when(
      err_reported == "nanas" ~ "nans",
      TRUE ~ err_reported
    )
  ) %>%
  # filter out 's'
  # 's' is the 'skip' option
  filter(err_reported_clean != "s") %>%

  filter(stim_type == "target") %>% 
  filter(condition.x %in% c("correct", "phon_err", "non_phon_err", "both_err")) %>%
  
  # now we only focus on 'y', 'nans', 'nun'
  # 'y' stands for 'yes, the question needs to be corrected'
  # 'nans' stands for 'the question does NOT need any corrections'
  # 'nun' stands for 'not sure / unsure'
  filter(err_reported_clean %in% c("y", "nans", "nun"))


#### Was there error reported? Grouped bar chart with error bars ####
df_analysis_err <- df_analysis_err %>%
  mutate(
    err_reported_label = case_when(
      err_reported_clean == "nans" ~ "Question has no errors",
      err_reported_clean == "nun"  ~ "Unsure",
      err_reported_clean == "y"    ~ "Question needs correction"
    ),
    condition.x = factor(condition.x, levels = c("correct", "phon_err", "non_phon_err", "both_err"))
  )

# calculate proportions per participant
participant_props <- df_analysis_err %>%
  # count responses for each participant/condition/response_type
  count(participant_id, condition.x, err_reported_label) %>%
  complete(participant_id, condition.x, err_reported_label, fill = list(n = 0)) %>%
  
  # calculate per-participant proportions
  group_by(participant_id, condition.x) %>%
  mutate(
    Total_Trials = sum(n),
    Proportion = n / Total_Trials
  ) %>%
  ungroup()

# calculate group mean and SD
plot_summary_with_error_grouped <- participant_props %>%
  group_by(condition.x, err_reported_label) %>%
  summarise(
    N = n(),
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = Mean_Proportion - SE,
    ymax = Mean_Proportion + SE
  )


##### group bar chart #####
p_grouped_chart_err_reported <- ggplot(plot_summary_with_error_grouped, 
                           aes(x = condition.x, 
                               y = Mean_Proportion, 
                               fill = err_reported_label,
                               ymin = ymin,
                               ymax = ymax
                           )) +
  
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black") +
  geom_errorbar(
    position = position_dodge(width = 0.9), 
    width = 0.25, 
    color = "black"
  ) +
  scale_fill_brewer(palette = "Set3") + 
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Condition",
    y = "Proportion of Responses",
    fill = "Participant Response"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top"
  )

print(p_grouped_chart_err_reported)

#### Did participants report the intended errors when they said that there were errors (needs correction)? Crosstab ####

participant_detections <- df_analysis_err %>%
  group_by(participant_id, condition.x) %>%
  summarise(
    # total number of trials for this participant/condition
    Total_Response = n(), 
    
    # total number of detection (err_reported_clean == "y")
    Total_Error_Reported = sum(err_reported_clean == "y"),
    
    # phon_err is detected if they said "p" OR "both" AND reported an error
    Phon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% c("p", "both")),
    
    # "non_phon_err" is detected if they said "np" OR "both" AND reported an error
    NonPhon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% c("np", "both")),
    
    .groups = "drop"
  ) %>%
  filter(condition.x != "correct") 


# calculate the detection proportion per participant

participant_proportions <- participant_detections %>%
  mutate(
    `phonological error` = Phon_Detected_Count / Total_Response,
    `nonphonological error` = NonPhon_Detected_Count / Total_Response
  ) %>%
  select(participant_id, Stimulus_Condition = condition.x, `phonological error`, `nonphonological error`) %>%
  pivot_longer(
    cols = c(`phonological error`, `nonphonological error`),
    names_to = "Error_Type_Detected",
    values_to = "Proportion"
  )

plot_summary_with_error <- participant_proportions %>%
  filter(
    (Stimulus_Condition == "phon_err" & Error_Type_Detected == "phonological error") |
    (Stimulus_Condition == "non_phon_err" & Error_Type_Detected == "nonphonological error") |
    (Stimulus_Condition == "both_err") 
  ) %>%
  group_by(Stimulus_Condition, Error_Type_Detected) %>%
  summarise(
    N = n(), 
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = Mean_Proportion - SE,
    ymax = Mean_Proportion + SE
  ) %>%
  mutate(Stimulus_Condition = factor(Stimulus_Condition, 
                                     levels = c("phon_err", "non_phon_err", "both_err")))

##### plot hit rate chart #####
p_hit_rate_chart <- ggplot(plot_summary_with_error, 
                           aes(x = Stimulus_Condition, 
                               y = Mean_Proportion,
                               fill = Error_Type_Detected,
                               ymin = ymin,
                               ymax = ymax
                           )) +
  
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", linewidth = 0.2) +
  
  geom_errorbar(
    position = position_dodge(width = 0.9), 
    width = 0.25,
    color = "black"
  ) +
  
  scale_fill_manual(values = c("phonological error" = "skyblue", 
                               "nonphonological error" = "salmon")) +
  
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Condition",
    y = "Proportion of Errors Detected",
    fill = "Error Type Detected"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top"
  )

print(p_hit_rate_chart)

#### Rate of phonological error detection rate over the experiment ####

# here we examine whether participants got better or worse at detecting phonological errors as the experiment progressed
# we'll focus only on trials where a phonological error was present (phon_err or both_err)

##### group phon err detection change #####
df_analysis_err <- df_analysis_err %>%
  group_by(participant_id) %>%
  mutate(trial_num_sequential = row_number()) %>%
  ungroup()

# filter data to only include trials where a PHONOLOGICAL error was present
df_phon_error_trials <- df_analysis_err %>%
  filter(condition.x %in% c("phon_err", "both_err"))

# calculate a binary variable for successful PHONOLOGICAL detection
# detection is: error reported ('y') AND a 'p' or 'both' response type
df_phon_detection_time <- df_phon_error_trials %>%
  mutate(
    Phonological_Detection_Success = as.integer(
      err_reported_clean == "y" & err_type_response %in% c("p", "both")
    )
  )

time_change_overall <- df_phon_detection_time %>%
  group_by(trial_num_sequential, condition.x) %>%
  summarise(
    N = n(),
    Mean_Detection = mean(Phonological_Detection_Success, na.rm = TRUE),
    SD = sd(Phonological_Detection_Success, na.rm = TRUE),
    SE = SD / sqrt(N),
    ymin = Mean_Detection - SE,
    ymax = Mean_Detection + SE,
    .groups = "drop"
  )


p_detection_over_time_overall <- ggplot(time_change_overall, 
                                        aes(x = trial_num_sequential, 
                                            y = Mean_Detection, 
                                            color = condition.x)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.7) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2) +
  geom_smooth(
    data = df_phon_detection_time, 
    aes(y = Phonological_Detection_Success, color = condition.x), 
    method = "glm", 
    method.args = list(family = "binomial"), 
    se = TRUE, 
    linewidth = 1
  ) +
  scale_y_continuous(limits = c(0, 1.0), labels = scales::percent) +
  scale_x_continuous(breaks = seq(0, max(time_change_overall$trial_num_sequential), by = 2)) +
  labs(
    x = "Trial Number",
    y = "Mean Proportion of Intended Phonological Errors Detected",
    color = "Error Condition"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top")

print(p_detection_over_time_overall)

##### participant level detection rate change #####

df_analysis_err <- df_analysis_err %>%
  group_by(participant_id) %>%
  mutate(trial_num_sequential = row_number()) %>%
  ungroup()

# filter data to only include trials where a PHONOLOGICAL error was present
df_phon_error_trials <- df_analysis_err %>%
  filter(condition.x %in% c("phon_err", "both_err"))

# calculate a binary variable for successful PHONOLOGICAL detection
# detection is: error reported ('y') AND a 'p' or 'both' response type
df_phon_detection_time <- df_phon_error_trials %>%
  mutate(
    Phonological_Detection_Success = as.integer(
      err_reported_clean == "y" & err_type_response %in% c("p", "both")
    )
  )

p_detection_per_participant <- ggplot(df_phon_detection_time, 
                                      aes(x = trial_num_sequential, 
                                          y = Phonological_Detection_Success, 
                                          color = condition.x)) +

  geom_point(position = position_jitter(height = 0.05), alpha = 0.5, size = 1.5) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE, 
              linewidth = 1) +
  facet_wrap(~ participant_id, ncol = 6) +
  scale_y_continuous(breaks = c(0, 1), 
                     labels = c("Miss", "Hit"), 
                     limits = c(-0.1, 1.1)) +
  
  labs(
    x = "Trial Number",
    y = "Phonological Error Detection Success",
    color = "Error Condition"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top"
  )

print(p_detection_per_participant)


#### Target sounds vs. Other sounds  ####

participant_response_targ_other_sound <- df_analysis_err %>%
  filter(condition.x %in% c("phon_err", "both_err")) %>%
  group_by(participant_id, condition.x) %>%
  summarise(
    Total_Response = n(), 
    P_Response_Count = sum(err_type_response == "p", na.rm = TRUE),
    OtherP_Response_Count = sum(err_type_response == "otherp", na.rm = TRUE),
    .groups = "drop"
  )

participant_proportions_targ_other_sound <- participant_response_targ_other_sound %>%
  mutate(
    `Phonological (p) Response` = P_Response_Count / Total_Response,
    `Other Phonological (otherp) Response` = OtherP_Response_Count / Total_Response
  ) %>%
  select(participant_id, Stimulus_Condition = condition.x, 
         `Phonological (p) Response`, `Other Phonological (otherp) Response`) %>%
  pivot_longer(
    cols = c(`Phonological (p) Response`, `Other Phonological (otherp) Response`),
    names_to = "Error_Classification_Reported",
    values_to = "Proportion"
  )


plot_summary_targ_other_sound <- participant_proportions_targ_other_sound %>%
  group_by(Stimulus_Condition, Error_Classification_Reported) %>%
  summarise(
    N = n(), 
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = Mean_Proportion - SE,
    ymax = Mean_Proportion + SE
  ) %>%
  mutate(Stimulus_Condition = factor(Stimulus_Condition, 
                                     levels = c("phon_err", "non_phon_err", "both_err")))


p_target_sound_reporting <- ggplot(plot_summary_targ_other_sound, 
                                   aes(x = Stimulus_Condition, 
                                       y = Mean_Proportion,
                                       fill = Error_Classification_Reported,
                                       ymin = ymin,
                                       ymax = ymax
                                   )) +
  
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", linewidth = 0.2) +
  
  geom_errorbar(
    position = position_dodge(width = 0.9), 
    width = 0.25,
    color = "black"
  ) +
  
  scale_fill_manual(values = c("Phonological (p) Response" = "skyblue", 
                               "Other Phonological (otherp) Response" = "lightcoral")) +
  
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, max(plot_summary_targ_other_sound$ymax) * 1.05)) +
  labs(
    x = "Error Condition",
    y = "Mean Proportion of Responses",
    fill = "Reported Error Type"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top"
  )

print(p_target_sound_reporting)


#### Role of accent familiarity ####

familiarity_levels <- c("Never", "Sometimes", "Half the time", "Often", "All the time")


df_familiarity <- df_final_joined %>%
  filter(stimuli_presented == "exposure_spanish_accent") %>%
  select(participant_id, response_value.y) %>%
  distinct() %>%
  rename(exposure_spanish_accent = response_value.y)

# join familiarity data with main analysis data
df_analysis_with_familiarity <- df_analysis_err %>%
  left_join(df_familiarity, by = "participant_id") %>%
  mutate(
    exposure_spanish_accent = factor(exposure_spanish_accent, 
                                     levels = familiarity_levels,
                                     ordered = TRUE)
  ) %>%
  filter(!is.na(exposure_spanish_accent)) 

# calculate detection proportions per participant and per familiarity group
participant_detections_fam <- df_analysis_with_familiarity %>%
  group_by(participant_id, exposure_spanish_accent, condition.x) %>%
  summarise(
    Total_Response = n(), 
    # Phon_Detected is a 'y' response AND type 'p' or 'both'
    Phon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% c("p", "both")),
    # NonPhon_Detected is a 'y' response AND type 'np' or 'both'
    NonPhon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% c("np", "both")),
    .groups = "drop"
  ) %>%
  filter(condition.x != "correct") 

# calculate proportion per participant
participant_proportions_fam <- participant_detections_fam %>%
  mutate(
    `phonological error` = Phon_Detected_Count / Total_Response,
    `nonphonological error` = NonPhon_Detected_Count / Total_Response
  ) %>%
  select(participant_id, exposure_spanish_accent, Stimulus_Condition = condition.x, 
         `phonological error`, `nonphonological error`) %>%
  pivot_longer(
    cols = c(`phonological error`, `nonphonological error`),
    names_to = "Error_Type_Detected",
    values_to = "Proportion"
  )

plot_summary_with_error_fam <- participant_proportions_fam %>%
  filter(
    (Stimulus_Condition == "phon_err" & Error_Type_Detected == "phonological error") |
    (Stimulus_Condition == "non_phon_err" & Error_Type_Detected == "nonphonological error") |
    (Stimulus_Condition == "both_err") 
  ) %>%
  group_by(exposure_spanish_accent, Stimulus_Condition, Error_Type_Detected) %>%
  summarise(
    N = n(), 
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = Mean_Proportion - SE,
    ymax = Mean_Proportion + SE
  ) %>%
  mutate(Stimulus_Condition = factor(Stimulus_Condition, 
                                     levels = c("phon_err", "non_phon_err", "both_err")))

##### accent familiarity X hit rate #####
p_hit_rate_chart_familiarity <- ggplot(plot_summary_with_error_fam, 
                                        aes(x = Stimulus_Condition, 
                                            y = Mean_Proportion, 
                                            fill = Error_Type_Detected,
                                            ymin = ymin,
                                            ymax = ymax)) +
  
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", linewidth = 0.2) +
  
  geom_errorbar(
    position = position_dodge(width = 0.9), 
    width = 0.25,
    color = "black"
  ) +
  
  scale_fill_manual(values = c("phonological error" = "skyblue", 
                               "nonphonological error" = "salmon")) +
  
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1.0)) +
  facet_wrap(~ exposure_spanish_accent, ncol = 3) +
  
  labs(
    x = "Condition",
    y = "Proportion of Errors Detected",
    fill = "Error Type Detected"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold") 
  )

print(p_hit_rate_chart_familiarity)




##### accent familiarity X detection of phon errors over time #####

df_detection_time_with_familiarity <- df_phon_detection_time %>%
  left_join(df_familiarity, by = "participant_id") %>%
  mutate(
    exposure_spanish_accent = factor(exposure_spanish_accent, 
                                     levels = familiarity_levels,
                                     ordered = TRUE)
  ) %>%
  filter(!is.na(exposure_spanish_accent)) 

time_change_by_fam <- df_detection_time_with_familiarity %>%
  group_by(trial_num_sequential, condition.x, exposure_spanish_accent) %>%
  summarise(
    N = n(),
    Mean_Detection = mean(Phonological_Detection_Success, na.rm = TRUE),
    SD = sd(Phonological_Detection_Success, na.rm = TRUE),
    SE = SD / sqrt(N),
    ymin = Mean_Detection - SE,
    ymax = Mean_Detection + SE,
    .groups = "drop"
  )

p_detection_time_by_familiarity <- ggplot(df_detection_time_with_familiarity, 
                                        aes(x = trial_num_sequential, 
                                            y = Phonological_Detection_Success, 
                                            color = condition.x)) +
  geom_point(position = position_jitter(height = 0.05), alpha = 0.15, size = 0.8) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE, 
              linewidth = 1) +
  facet_wrap(~ exposure_spanish_accent, ncol = 3) +
  
  scale_y_continuous(breaks = c(0, 1), 
                     labels = c("Miss", "Hit"), 
                     limits = c(-0.1, 1.1)) +
  
  labs(
    x = "Trial Number",
    y = "Phonological Error Detection Success",
    color = "Error Condition"
  ) +
  theme_bw() +
  theme(
   axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold", size = 14),
    legend.text = element_text(size = 14),
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p_detection_time_by_familiarity)

```


#### 2.2 Main batch3: 3 subjects data (both attempts)


```{r error-detection-annotated-data-main-batch3-3subj, message=FALSE, warning=FALSE}

# load data and exclude participants who failed the attention checks


tryCatch({
 
  df_fb <- read_csv("../data/annotated/main_batch3_test_run_FB.csv")
  df_jb <- read_csv("../data/annotated/main_batch3_test_run_JB.csv")
  df_quest <- read_csv("../data/preprocessed/main_batch3/testrun_4subj_questionnaire_data.csv")
  
  df_exp_all <- bind_rows(df_fb, df_jb)
  
  # remove participants who are completely missing annotations (no coder ID found in any of their rows)
  # this means that the participants was not engaging with the task
  df_exp_all <- df_exp_all %>%
    group_by(participant_id) %>%
    filter(any(!is.na(coder))) %>%
    ungroup()

  # attention checks (<6/8 OR 4/7)
  # one misannotated 4/7 in comments
  
  df_exp_all <- df_exp_all %>%
    mutate(
      
      comments = as.character(comments),
      
     
      attention_score_8 = as.numeric(
        str_extract(comments, "(?<=\\()\\d+(?=\\/8)")
      ),
      
     
      attention_score_7 = as.numeric(
        str_extract(comments, "(?<=\\()\\d+(?=\\/7)")
      )
    )


  failed_participants_8 <- df_exp_all %>%
    filter(attention_score_8 < 6) %>%
    pull(participant_id) %>%
    unique()
    
  failed_participants_7 <- df_exp_all %>%
    filter(attention_score_7 == 4) %>%
    pull(participant_id) %>%
    unique()
  
 
  failed_participants <- union(failed_participants_8, failed_participants_7)

  print(paste("Found", length(failed_participants), "total unique participants who failed."))
  print("Failed participant IDs:")
  print(failed_participants)


  df_exp_all <- df_exp_all %>%
    mutate(attention_check_passed = !participant_id %in% failed_participants)
  
 
# sanity checks after excluding those who failed the attention checks
  
  status_counts <- df_exp_all %>%
    distinct(participant_id, attention_check_passed) %>%
    count(attention_check_passed)
  
  print(status_counts)
 

# join combined annotated, cleaned error detection data with questionnaire data
  df_final_joined <- df_exp_all %>%
    left_join(df_quest, by = "participant_id")
})


df_clean_attchp <- df_final_joined %>%
  filter(attention_check_passed == TRUE) %>% 
  distinct(participant_id, question, .keep_all = TRUE)

print(paste("Analyzing", n_distinct(df_clean_attchp$participant_id), "participants who passed attention checks."))
```


```{r error-detection-data-plotting-main-batch3-3subj, message=FALSE, warning=FALSE}

phon_hit_types <- c("p", "p?", "both")
nonphon_hit_types <- c("np", "np?", "both")

df_analysis_err <- df_clean_attchp %>%
  mutate(
    err_reported_clean = case_when(
      err_reported == "nanas" ~ "nans",
      TRUE ~ err_reported
    )
  ) %>%
  filter(err_reported_clean != "s") %>% 
  filter(stim_type == "target") %>% 
  filter(condition.x %in% c("correct", "phon_err", "non_phon_err", "both_err")) %>%
  filter(err_reported_clean %in% c("y", "nans", "nun"))

df_analysis_err <- df_analysis_err %>%
  mutate(
    condition.x = factor(condition.x, levels = c("correct", "phon_err", "non_phon_err", "both_err")),
    err_reported_label = case_when(
      err_reported_clean == "nans" ~ "Question has no errors",
      err_reported_clean == "nun"  ~ "Unsure",
      err_reported_clean == "y"    ~ "Question needs correction"
    )
  )


#### (A) Was there error reported? ####

# % per participant
participant_props <- df_analysis_err %>%
  count(participant_id, condition.x, err_reported_label) %>%
  complete(participant_id, condition.x, err_reported_label, fill = list(n = 0)) %>%
  group_by(participant_id, condition.x) %>%
  mutate(
    Total_Trials = sum(n),
    Proportion = n / Total_Trials
  ) %>%
  ungroup()

plot_summary_err_reported <- participant_props %>%
  group_by(condition.x, err_reported_label) %>%
  summarise(
    N = n(),
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(ymin = Mean_Proportion - SE, ymax = Mean_Proportion + SE)

p_grouped_chart_err_reported <- ggplot(plot_summary_err_reported, 
                                     aes(x = condition.x, y = Mean_Proportion, fill = err_reported_label, ymin = ymin, ymax = ymax)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black") +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_brewer(palette = "Set3") + 
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Condition", y = "Proportion of Responses", fill = "Participant Response", title = "Did participants report an error?") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), legend.position = "top")

print(p_grouped_chart_err_reported)


#### (B) Hit Rate: Did they report the INTENDED error? ####

participant_detections <- df_analysis_err %>%
  group_by(participant_id, condition.x) %>%
  summarise(
    Total_Response = n(),
    Phon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% phon_hit_types),
    NonPhon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% nonphon_hit_types),
    
    .groups = "drop"
  ) %>%
  filter(condition.x != "correct") 

participant_proportions <- participant_detections %>%
  mutate(
    `phonological error` = Phon_Detected_Count / Total_Response,
    `nonphonological error` = NonPhon_Detected_Count / Total_Response
  ) %>%
  select(participant_id, Stimulus_Condition = condition.x, `phonological error`, `nonphonological error`) %>%
  pivot_longer(cols = c(`phonological error`, `nonphonological error`), names_to = "Error_Type_Detected", values_to = "Proportion")

plot_summary_hits <- participant_proportions %>%
  filter(
    (Stimulus_Condition == "phon_err" & Error_Type_Detected == "phonological error") |
    (Stimulus_Condition == "non_phon_err" & Error_Type_Detected == "nonphonological error") |
    (Stimulus_Condition == "both_err") 
  ) %>%
  group_by(Stimulus_Condition, Error_Type_Detected) %>%
  summarise(
    N = n(), Mean_Proportion = mean(Proportion, na.rm = TRUE), SD = sd(Proportion, na.rm = TRUE), SE = SD / sqrt(N), .groups = "drop"
  ) %>%
  mutate(ymin = Mean_Proportion - SE, ymax = Mean_Proportion + SE)

p_hit_rate_chart <- ggplot(plot_summary_hits, 
                           aes(x = Stimulus_Condition, y = Mean_Proportion, fill = Error_Type_Detected, ymin = ymin, ymax = ymax)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black") +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("phonological error" = "skyblue", "nonphonological error" = "salmon")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Condition", y = "Detection Rate", fill = "Error Type Detected", title = "Hit Rate: Detection of Intended Errors") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), legend.position = "top")

print(p_hit_rate_chart)

#### (C) When was the error reported? (err_reported_attempt) ####
# definition: Reporting for the first or second attempt?
# filter: Only look at cases where an error WAS reported (y)
# categories: first, secsame, secdiff

df_attempt_analysis <- df_analysis_err %>%
  filter(err_reported_clean == "y") %>%
  filter(!is.na(err_reported_attempt)) %>%
  mutate(
    attempt_label = case_when(
      err_reported_attempt == "first" ~ "First Attempt",
      err_reported_attempt == "secsame" ~ "Second Attempt (Same/Skip)",
      err_reported_attempt == "secdiff" ~ "Second Attempt (Changed)",
      TRUE ~ err_reported_attempt
    )
  )

# % of attempt types within the "Error Reported" trials
plot_summary_attempt <- df_attempt_analysis %>%
  group_by(condition.x, attempt_label) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(condition.x) %>%
  mutate(Proportion = Count / sum(Count))

p_attempt_chart <- ggplot(plot_summary_attempt, 
                          aes(x = condition.x, y = Proportion, fill = attempt_label)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Pastel1") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Condition", 
    y = "Proportion of Reported Errors", 
    fill = "Reporting Attempt",
    title = "When did participants report the error?"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), legend.position = "top")

print(p_attempt_chart)


#### (D) Accuracy of "No Error" Responses (no_err_response) ####
# definition: In reportedly no-error questions, are they accurate?
# filter: Only look at cases where NO error was reported (nans)
# categories: a (accurate), w (wrong), blank, corrected, twice

df_noerr_analysis <- df_analysis_err %>%
  filter(err_reported_clean == "nans") %>%
  filter(!is.na(no_err_response)) %>%
  mutate(
    accuracy_label = case_when(
      no_err_response == "a" ~ "Accurate",
      no_err_response == "w" ~ "Wrong",
      no_err_response == "blank" ~ "Blank",
      no_err_response == "corrected" ~ "Corrected (1st wrong -> 2nd right)",
      no_err_response == "twice" ~ "Same Response Twice", # Usually implies consistent answer
      TRUE ~ no_err_response
    )
  )

plot_summary_noerr <- df_noerr_analysis %>%
  group_by(condition.x, accuracy_label) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(condition.x) %>%
  mutate(Proportion = Count / sum(Count))

p_noerr_accuracy_chart <- ggplot(plot_summary_noerr, 
                                 aes(x = condition.x, y = Proportion, fill = accuracy_label)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Condition", 
    y = "Proportion of 'No Error' Responses", 
    fill = "Comprehension Accuracy",
    title = "Accuracy of trivia question answers when NO error was reported"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), legend.position = "top")

print(p_noerr_accuracy_chart)


#### (E) Detailed Error Type Breakdown ####
# what kind of errors are being reported in detail

df_detailed_types <- df_analysis_err %>%
  filter(err_reported_clean == "y") %>%
  count(condition.x, err_type_response) %>%
  group_by(condition.x) %>%
  mutate(Proportion = n / sum(n))

p_detailed_types <- ggplot(df_detailed_types, aes(x = err_type_response, y = Proportion, fill = condition.x)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  facet_wrap(~condition.x, scales = "free_x") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Error Type Code", 
    y = "Proportion of Reported Errors",
    title = "Detailed Breakdown of Reported Error Types"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10), legend.position = "none")

print(p_detailed_types)


#### (F) Rate of phonological error detection rate over the experiment ####

df_analysis_err <- df_analysis_err %>%
  group_by(participant_id) %>%
  mutate(trial_num_sequential = row_number()) %>%
  ungroup()

df_phon_error_trials <- df_analysis_err %>%
  filter(condition.x %in% c("phon_err", "both_err"))

df_phon_detection_time <- df_phon_error_trials %>%
  mutate(
    Phonological_Detection_Success = as.integer(
      err_reported_clean == "y" & err_type_response %in% phon_hit_types
    )
  )

p_detection_over_time_overall <- ggplot(df_phon_detection_time, 
                                        aes(x = trial_num_sequential, y = Phonological_Detection_Success, color = condition.x)) +
  geom_point(position = position_jitter(height = 0.05), alpha = 0.3) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, linewidth = 1) +
  scale_y_continuous(breaks = c(0, 1), labels = c("Miss", "Hit"), limits = c(-0.1, 1.1)) +
  labs(x = "Trial Number", y = "Phonological Error Detection Success", color = "Condition", title = "Phonological error detection rate over time") +
  theme_bw()

print(p_detection_over_time_overall)


#### (G) Role of Spanish accent familiarity ####

familiarity_levels <- c("Never", "Sometimes", "Half the time", "Often", "All the time")


df_familiarity <- df_final_joined %>%
  filter(stimuli_presented == "exposure_spanish_accent") %>%
  select(participant_id, response_value.y) %>%
  distinct() %>%
  rename(exposure_spanish_accent = response_value.y)


df_analysis_with_familiarity <- df_analysis_err %>%
  left_join(df_familiarity, by = "participant_id") %>%
  mutate(
    exposure_spanish_accent = factor(
      exposure_spanish_accent,
      levels = c(1, 2, 3, 4, 5),      
      labels = familiarity_levels,   
      ordered = TRUE
    )
  ) %>%
  filter(!is.na(exposure_spanish_accent))

# detection proportions per participant and per familiarity group
participant_detections_fam <- df_analysis_with_familiarity %>%
  group_by(participant_id, exposure_spanish_accent, condition.x) %>%
  summarise(
    Total_Response = n(), 
    Phon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% phon_hit_types),
    NonPhon_Detected_Count = sum(err_reported_clean == "y" & err_type_response %in% nonphon_hit_types),
    .groups = "drop"
  ) %>%
  filter(condition.x != "correct") 

# proportion per participant
participant_proportions_fam <- participant_detections_fam %>%
  mutate(
    `phonological error` = Phon_Detected_Count / Total_Response,
    `nonphonological error` = NonPhon_Detected_Count / Total_Response
  ) %>%
  select(participant_id, exposure_spanish_accent, Stimulus_Condition = condition.x, 
         `phonological error`, `nonphonological error`) %>%
  pivot_longer(
    cols = c(`phonological error`, `nonphonological error`),
    names_to = "Error_Type_Detected",
    values_to = "Proportion"
  )

plot_summary_with_error_fam <- participant_proportions_fam %>%
  filter(
    (Stimulus_Condition == "phon_err" & Error_Type_Detected == "phonological error") |
    (Stimulus_Condition == "non_phon_err" & Error_Type_Detected == "nonphonological error") |
    (Stimulus_Condition == "both_err") 
  ) %>%
  group_by(exposure_spanish_accent, Stimulus_Condition, Error_Type_Detected) %>%
  summarise(
    N = n(), 
    Mean_Proportion = mean(Proportion, na.rm = TRUE),
    SD = sd(Proportion, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(
    ymin = Mean_Proportion - SE,
    ymax = Mean_Proportion + SE
  ) %>%
  mutate(Stimulus_Condition = factor(Stimulus_Condition, levels = c("phon_err", "non_phon_err", "both_err")))



##### (G1) Accent familiarity X Hit rate #####
p_hit_rate_chart_familiarity <- ggplot(plot_summary_with_error_fam, 
                                       aes(x = Stimulus_Condition, 
                                           y = Mean_Proportion, 
                                           fill = Error_Type_Detected, 
                                           ymin = ymin, 
                                           ymax = ymax)) +
  
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), color = "black", linewidth = 0.2) +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("phonological error" = "skyblue", "nonphonological error" = "salmon")) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1.0)) +
  facet_wrap(~ exposure_spanish_accent, ncol = 3) +
  labs(x = "Condition", y = "Proportion of Errors Detected", fill = "Error Type Detected", title = "Hit Rate by Accent Familiarity") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold") 
  )

print(p_hit_rate_chart_familiarity)


##### (G2) Accent familiarity X Detection of phon errors over time #####

familiarity_levels <- c("Never", "Sometimes", "Half the time", "Often", "All the time")

df_detection_time_with_familiarity <- df_phon_detection_time %>%
  left_join(df_familiarity, by = "participant_id") %>%
  mutate(
    exposure_spanish_accent = factor(exposure_spanish_accent, 
                                     levels = c(1, 2, 3, 4, 5),   
                                     labels = familiarity_levels,
                                     ordered = TRUE)
  ) %>%
  filter(!is.na(exposure_spanish_accent)) 

p_detection_time_by_familiarity <- ggplot(df_detection_time_with_familiarity, 
                                      aes(x = trial_num_sequential, 
                                          y = Phonological_Detection_Success, 
                                          color = condition.x)) +
  geom_point(position = position_jitter(height = 0.05), alpha = 0.15, size = 0.8) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = TRUE, 
              linewidth = 1) +
  facet_wrap(~ exposure_spanish_accent) +
  scale_y_continuous(breaks = c(0, 1), labels = c("Miss", "Hit"), limits = c(-0.1, 1.1)) +
  labs(x = "Trial Number", 
       y = "Phonological Error Detection Success", 
       color = "Error Condition", 
       title = "Phonological Detection Over Time by Familiarity") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    legend.position = "top",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p_detection_time_by_familiarity)


#### (H) False Alarm (Response Bias) ####
# definition: How often do participants report an error when the sentence is actually CORRECT?

df_false_alarms <- df_analysis_err %>%
  filter(condition.x == "correct") %>%
  group_by(participant_id) %>%
  summarise(
    Total_Correct_Trials = n(),
    False_Alarms = sum(err_reported_clean == "y"),
    FA_Rate = False_Alarms / Total_Correct_Trials,
    .groups = "drop"
  )

p_false_alarms <- ggplot(df_false_alarms, aes(x = FA_Rate)) +
  geom_histogram(binwidth = 0.05, fill = "salmon", color = "black", alpha = 0.7) +
  geom_vline(aes(xintercept = mean(FA_Rate, na.rm=TRUE)), color="red", linetype="dashed", linewidth=1) +
  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = "False Alarm Rate (Proportion)", 
       y = "Count of Participants",
       title = "How often do participants report an error when the sentence is actually CORRECT?",
       subtitle = "Dashed line = Mean False Alarm Rate") +
  theme_bw()

print(p_false_alarms)


#### (I) Interference Analysis ####
##### (I1)Does the presence of a Non-Phonological error make it harder to detect the Phonological one? #####
# compare 'phon_err' (Single) vs 'both_err' (Double)

df_interference <- df_analysis_err %>%
  filter(condition.x %in% c("phon_err", "both_err")) %>%
  mutate(
    detected_phon = as.integer(err_reported_clean == "y" & err_type_response %in% phon_hit_types)
  )

p_interference <- ggplot(df_interference, aes(x = condition.x, y = detected_phon, fill = condition.x)) +
  stat_summary(fun = "mean", geom = "bar", color = "black", width = 0.6) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
  # Add individual points to show spread
  stat_summary(fun = "mean", geom = "point", shape = 18, size = 3) +
  scale_fill_brewer(palette = "Pastel2") +
  labs(x = "Condition", 
       y = "Phonological Error Detection Rate",
       title = "Does a non-sound error mask the sound error?")+
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_bw() + 
  theme(legend.position = "none")

print(p_interference)


##### (I2)Does the presence of a Phonological error make it harder to detect the Non-Phonological one? #####

# compare "non_phon_err" vs "both_err" 
# check for detection of the non-phon error
df_interference_inverse <- df_analysis_err %>%
  filter(condition.x %in% c("non_phon_err", "both_err")) %>%
  mutate(
    detected_nonphon = as.integer(err_reported_clean == "y" & err_type_response %in% nonphon_hit_types)
  )

p_interference_inverse <- ggplot(df_interference_inverse, 
                                 aes(x = condition.x, y = detected_nonphon, fill = condition.x)) +
  stat_summary(fun = "mean", geom = "bar", color = "black", width = 0.6) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2) +
  stat_summary(fun = "mean", geom = "point", shape = 18, size = 3) +
  scale_fill_brewer(palette = "Pastel1") + # Different palette to distinguish from previous plot
  labs(x = "Condition", 
       y = "Non-Phonological Error Detection Rate",
       title = "Does a sound error affect detection of the non-sound error?") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_bw() + 
  theme(legend.position = "none")

print(p_interference_inverse)

```




