---
title: "Experiment 1: Error Detection in Spanish-Accented English"
subtitle: "Comprehensive Main Analysis"
author: "Yuting Gu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)

library(tidyverse)
library(lme4)
library(lmerTest)
library(scales)
library(rstatix)
library(ggpubr)
library(patchwork)
library(Hmisc)
library(tidytext)

# Color palette
COL_PHON    <- "#E07B3F"
COL_NONPHON <- "#2B7BB9"
COL_CORRECT <- "#57A75C"
COL_BOTH    <- "#9B6EBD"

error_colors <- c("Phonological" = COL_PHON, "Non-Phonological" = COL_NONPHON)

# Poster-friendly theme
theme_set(theme_bw(base_size = 16) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill = "grey95"),
    legend.position = "top"
  ))

# Formatting helpers
fmt_p <- function(p) {
  if (p < .001) "p < .001"
  else if (p < .01) sprintf("p = %.3f", p)
  else sprintf("p = %.2f", p)
}

extract_pval <- function(model, term) {
  coefs <- summary(model)$coefficients
  if (term %in% rownames(coefs)) coefs[term, "Pr(>|z|)"]
  else NA_real_
}
```


# 1. Data Loading & Exclusions

```{r data-loading}
process_batch_error_detection <- function(file_prefix, group_label,
                                           batch_label, has_questionnaire) {
  annotated_dir <- "../data/annotated"

  if (!dir.exists(annotated_dir)) {
    stop(sprintf("Directory not found: %s\nWorking directory: %s",
                 normalizePath(annotated_dir, mustWork = FALSE), getwd()))
  }

  # Find CSVs matching the prefix (e.g., main_batch3_FB.csv) but exclude test runs
  csvs <- list.files(annotated_dir,
                      pattern = paste0("^", file_prefix, "_.*\\.csv$"),
                      full.names = TRUE)
  csvs <- csvs[!str_detect(basename(csvs), "test_run")]

  if (length(csvs) == 0) {
    stop(sprintf("No CSV files matching '%s_*.csv' in %s",
                 file_prefix, normalizePath(annotated_dir, mustWork = FALSE)))
  }
  cat(sprintf("Found %d CSV files: %s\n", length(csvs), paste(basename(csvs), collapse = ", ")))

  d <- map_dfr(csvs, function(f) {
    read_csv(f, show_col_types = FALSE) %>%
      mutate(across(everything(), as.character))
  })

  # Extract attention check scores from comments
  att_df <- d %>%
    filter(str_detect(comments, "\\d+/8")) %>%
    mutate(att_score = as.integer(str_extract(comments, "\\d+(?=/8)"))) %>%
    select(participant_id, att_score) %>%
    distinct(participant_id, .keep_all = TRUE)

  failed_ids <- att_df %>% filter(att_score < 6) %>% pull(participant_id)
  message(sprintf("Batch '%s': %d participants, %d failed attention check (threshold 6/8)",
                  batch_label, n_distinct(d$participant_id), length(failed_ids)))

  d_clean <- d %>% filter(!participant_id %in% failed_ids)

  # Add questionnaire data if available
  if (has_questionnaire) {
    preproc_dir <- file.path("../data/preprocessed", file_prefix)
    q_search_dirs <- c(annotated_dir, preproc_dir)
    q_files <- unlist(lapply(q_search_dirs[dir.exists(q_search_dirs)],
                              list.files, pattern = "questionnaire_data.csv",
                              full.names = TRUE, recursive = TRUE))
    if (length(q_files) > 0) {
      df_q <- map_dfr(q_files, function(f) {
        read_csv(f, show_col_types = FALSE) %>%
          filter(stimuli_presented %in% c("exposure_spanish_accent", "exposure_spanish")) %>%
          select(participant_id, question_id = stimuli_presented, answer = response_value) %>%
          distinct() %>%
          pivot_wider(names_from = question_id, values_from = answer) %>%
          rename_with(~ paste0("quest_", .), .cols = -participant_id)
      }) %>%
        group_by(participant_id) %>% slice(1) %>% ungroup()

      d_clean <- left_join(d_clean, df_q, by = "participant_id")
    }
  }

  d_clean$batch_id <- batch_label
  d_clean$group    <- group_label
  list(data = d_clean, failed_ids = failed_ids, att_scores = att_df)
}

# Load main batch 3 data
batch_result <- process_batch_error_detection(
  "main_batch3", "main", "Main Batch 3", TRUE
)
df_all   <- batch_result$data
att_info <- batch_result$att_scores
```

```{r build-trial-df}
# Build trial-level data frame from annotated rows
# Each (participant_id, question) group represents one trial with multiple rows

df_trials <- df_all %>%
  filter(!is.na(stim_type), stim_type != "NA") %>%
  group_by(participant_id, question) %>%
  summarise(
    stim_type = first(na.omit(stim_type[stim_type != "NA"])),
    condition = first(na.omit(condition[condition != "NA"])),

    # First attempt
    err_reported_first = {
      idx <- which(err_reported_attempt == "first")
      if (length(idx) > 0) err_reported[idx[1]] else NA_character_
    },
    err_type_first = {
      idx <- which(err_reported_attempt == "first")
      if (length(idx) > 0) err_type_response[idx[1]] else NA_character_
    },

    # Second attempt
    err_reported_second = {
      idx <- which(str_starts(err_reported_attempt, "sec"))
      if (length(idx) > 0) err_reported[idx[1]] else NA_character_
    },
    err_type_second = {
      idx <- which(str_starts(err_reported_attempt, "sec"))
      if (length(idx) > 0) err_type_response[idx[1]] else NA_character_
    },
    second_attempt_used = {
      idx <- which(str_starts(err_reported_attempt, "sec"))
      if (length(idx) > 0) {
        attempt_val <- err_reported_attempt[idx[1]]
        resp_val    <- response_value[idx[1]] %||% ""
        # secdiff = changed answer, secsame = kept answer
        # Check if they actually re-listened vs. skipped
        !(str_detect(tolower(resp_val), "good") & str_detect(tolower(resp_val), "skip"))
      } else FALSE
    },
    .groups = "drop"
  ) %>%
  filter(stim_type == "target") %>%
  mutate(
    # Detection indicators for first attempt
    phon_detected_first = err_reported_first == "y" &
      err_type_first %in% c("p", "p?", "otherp", "both", "both?"),
    nonphon_detected_first = err_reported_first == "y" &
      err_type_first %in% c("np", "np?", "othernp", "both", "both?"),

    # Detection indicators for combined (first + second)
    phon_detected_comb = phon_detected_first | (
      err_reported_second == "y" &
        err_type_second %in% c("p", "p?", "otherp", "both", "both?")
    ),
    nonphon_detected_comb = nonphon_detected_first | (
      err_reported_second == "y" &
        err_type_second %in% c("np", "np?", "othernp", "both", "both?")
    )
  )

# Add trial position (presentation order within each participant)
df_trials <- df_trials %>%
  group_by(participant_id) %>%
  mutate(trial_position = row_number()) %>%
  ungroup()

cat(sprintf("Target trials: %d from %d participants\n",
            nrow(df_trials), n_distinct(df_trials$participant_id)))
cat(sprintf("Conditions: %s\n",
            paste(names(table(df_trials$condition)), collapse = ", ")))
```


# 2. Detection Rate Analysis (GLMER)

```{r build-long-df}
build_long_df <- function(df, attempt = "first") {
  # Create long-format data for GLMER: 2 rows per trial (Phon + NonPhon detection)

  detect_phon_col <- if (attempt == "first") "phon_detected_first" else "phon_detected_comb"
  detect_np_col   <- if (attempt == "first") "nonphon_detected_first" else "nonphon_detected_comb"

  phon_rows <- df %>%
    transmute(
      participant_id,
      question,
      condition,
      ErrorType = "Phonological",
      Present   = as.integer(condition %in% c("phon_err", "both_err")),
      OtherTypePresent = as.integer(condition %in% c("non_phon_err", "both_err")),
      Reported  = as.integer(!!sym(detect_phon_col))
    )

  np_rows <- df %>%
    transmute(
      participant_id,
      question,
      condition,
      ErrorType = "Non-Phonological",
      Present   = as.integer(condition %in% c("non_phon_err", "both_err")),
      OtherTypePresent = as.integer(condition %in% c("phon_err", "both_err")),
      Reported  = as.integer(!!sym(detect_np_col))
    )

  bind_rows(phon_rows, np_rows) %>%
    mutate(
      # Sum-coded ErrorType: NonPhon = +1, Phon = -1
      ErrorType1 = if_else(ErrorType == "Non-Phonological", 1, -1),
      Present_c  = Present - mean(Present),
      OtherTypePresent_c = OtherTypePresent - mean(OtherTypePresent)
    )
}

df_long_att1 <- build_long_df(df_trials, "first")
df_long_comb <- build_long_df(df_trials, "combined")
```

```{r glmer-model}
# Main GLMER model (first attempt)
m_glmer <- glmer(
  Reported ~ ErrorType1 * Present_c + ErrorType1 * OtherTypePresent_c +
    (1 | participant_id) + (1 | question),
  data   = df_long_att1,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

cat("GLMER Model: Reported ~ ErrorType * Present + ErrorType * OtherTypePresent\n")
cat("(sum-coded ErrorType: NonPhon = +1, Phon = -1)\n\n")
summary(m_glmer)
```

```{r detection-rates}
# Detection rates by condition (first attempt)
det_rates <- df_trials %>%
  summarise(
    phon_in_phon_err = mean(phon_detected_first[condition == "phon_err"]),
    phon_in_both_err = mean(phon_detected_first[condition == "both_err"]),
    nonphon_in_nonphon_err = mean(nonphon_detected_first[condition == "non_phon_err"]),
    nonphon_in_both_err = mean(nonphon_detected_first[condition == "both_err"]),
    fa_phon_in_correct = mean(phon_detected_first[condition == "correct"]),
    fa_nonphon_in_correct = mean(nonphon_detected_first[condition == "correct"]),
    fa_phon_in_nonphon = mean(phon_detected_first[condition == "non_phon_err"]),
    fa_nonphon_in_phon = mean(nonphon_detected_first[condition == "phon_err"])
  )

cat("Detection rates (first attempt):\n")
cat(sprintf("  Phon in phon_err:      %.3f\n", det_rates$phon_in_phon_err))
cat(sprintf("  Phon in both_err:      %.3f\n", det_rates$phon_in_both_err))
cat(sprintf("  NonPhon in nonphon_err:%.3f\n", det_rates$nonphon_in_nonphon_err))
cat(sprintf("  NonPhon in both_err:   %.3f\n", det_rates$nonphon_in_both_err))
cat(sprintf("  FA Phon in correct:    %.3f\n", det_rates$fa_phon_in_correct))
cat(sprintf("  FA NonPhon in correct: %.3f\n", det_rates$fa_nonphon_in_correct))
```


# 3. Signal Detection Theory Analysis

## 3.1 Standard SDT (d' and c by ErrorType)

```{r standard-sdt}
# Compute d' and c for each participant × ErrorType
# Signal = Present, Noise = Absent (collapsing across OtherTypePresent)

calc_sdt_standard <- function(df_long) {
  df_long %>%
    group_by(participant_id, ErrorType) %>%
    summarise(
      hits   = sum(Present == 1 & Reported == 1),
      misses = sum(Present == 1 & Reported == 0),
      fas    = sum(Present == 0 & Reported == 1),
      crs    = sum(Present == 0 & Reported == 0),
      n_signal = hits + misses,
      n_noise  = fas + crs,
      .groups = "drop"
    ) %>%
    mutate(
      # Log-linear correction (Hautus, 1995)
      hit_rate = (hits + 0.5) / (n_signal + 1),
      fa_rate  = (fas + 0.5)  / (n_noise + 1),
      d_prime  = qnorm(hit_rate) - qnorm(fa_rate),
      c_bias   = -0.5 * (qnorm(hit_rate) + qnorm(fa_rate))
    )
}

sdt_att1 <- calc_sdt_standard(df_long_att1) %>% mutate(Attempt = "Attempt 1")
sdt_comb <- calc_sdt_standard(df_long_comb) %>% mutate(Attempt = "Combined (Attempts 1 & 2)")
master_sdt <- bind_rows(sdt_att1, sdt_comb)

# Summary
master_sdt %>%
  group_by(Attempt, ErrorType) %>%
  summarise(
    d_M = mean(d_prime), d_SD = sd(d_prime),
    c_M = mean(c_bias),  c_SD = sd(c_bias),
    .groups = "drop"
  ) %>%
  knitr::kable(digits = 3, caption = "Standard SDT: d' and c by ErrorType")
```

```{r sdt-standard-tests}
# GLMER on d' and c (first attempt)
m1_att1 <- glmer(
  Reported ~ ErrorType1 + (1 | participant_id) + (1 | question),
  data    = df_long_att1,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

# Combined
m1_comb <- glmer(
  Reported ~ ErrorType1 + (1 | participant_id) + (1 | question),
  data    = df_long_comb,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

cat("ErrorType effect on detection (sum-coded, NonPhon=+1, Phon=-1):\n")
cat(sprintf("  Attempt 1:  p = %.4f\n", extract_pval(m1_att1, "ErrorType1")))
cat(sprintf("  Combined:   p = %.4f\n", extract_pval(m1_comb, "ErrorType1")))
```


## 3.2 Revised SDT: Competition Decomposition (d' and c by OtherTypePresent)

This analysis decomposes the competition (masking) effect into sensitivity loss vs. criterion shift by computing d' and c separately for single-error and both-error contexts.

Cell mapping for hit/FA computation:

- Phon detection, single (OTP=0): signal = phon_err, noise = correct
- Phon detection, both (OTP=1): signal = both_err, noise = non_phon_err
- NonPhon detection, single (OTP=0): signal = non_phon_err, noise = correct
- NonPhon detection, both (OTP=1): signal = both_err, noise = phon_err

Predictions:

- Resource-competition: d' drops when OTP=1, c unchanged
- Criterion shift / bias: c shifts when OTP=1, d' unchanged

```{r sdt-decomposition}
calc_sdt_decomposed <- function(df_long) {
  # Compute d' and c for each participant x ErrorType x OtherTypePresent
  # Using context-matched signal/noise conditions

  df_long %>%
    group_by(participant_id, ErrorType, OtherTypePresent) %>%
    summarise(
      hits   = sum(Present == 1 & Reported == 1),
      misses = sum(Present == 1 & Reported == 0),
      fas    = sum(Present == 0 & Reported == 1),
      crs    = sum(Present == 0 & Reported == 0),
      n_signal = hits + misses,
      n_noise  = fas + crs,
      .groups = "drop"
    ) %>%
    mutate(
      hit_rate_raw = if_else(n_signal > 0, hits / n_signal, NA_real_),
      fa_rate_raw  = if_else(n_noise > 0, fas / n_noise, NA_real_),
      # Log-linear correction (Hautus, 1995)
      hit_rate = (hits + 0.5) / (n_signal + 1),
      fa_rate  = (fas + 0.5)  / (n_noise + 1),
      d_prime  = qnorm(hit_rate) - qnorm(fa_rate),
      c_bias   = -0.5 * (qnorm(hit_rate) + qnorm(fa_rate)),
      Context  = if_else(OtherTypePresent == 0, "Single", "Both")
    )
}

sdt_decomp <- calc_sdt_decomposed(df_long_att1)

# Cell means table
sdt_decomp %>%
  group_by(ErrorType, Context) %>%
  summarise(
    d_M = mean(d_prime), d_SD = sd(d_prime),
    c_M = mean(c_bias),  c_SD = sd(c_bias),
    HR  = mean(hit_rate_raw, na.rm = TRUE),
    FAR = mean(fa_rate_raw, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  knitr::kable(digits = 3, caption = "SDT Decomposition: d' and c by ErrorType x Context")
```

```{r sdt-decomp-tests}
# 2x2 repeated-measures tests using paired t-tests

# Reshape to wide for paired tests
sdt_wide <- sdt_decomp %>%
  select(participant_id, ErrorType, Context, d_prime, c_bias) %>%
  pivot_wider(
    names_from  = c(ErrorType, Context),
    values_from = c(d_prime, c_bias),
    names_sep   = "_"
  )

# Main effect of OtherTypePresent on d'
sdt_wide <- sdt_wide %>%
  mutate(
    d_single_avg = (d_prime_Phonological_Single + `d_prime_Non-Phonological_Single`) / 2,
    d_both_avg   = (d_prime_Phonological_Both + `d_prime_Non-Phonological_Both`) / 2,
    c_single_avg = (c_bias_Phonological_Single + `c_bias_Non-Phonological_Single`) / 2,
    c_both_avg   = (c_bias_Phonological_Both + `c_bias_Non-Phonological_Both`) / 2
  )

cat("SENSITIVITY (d') TESTS\n\n")

# Main effect of OTP on d'
t_otp_d <- t.test(sdt_wide$d_single_avg, sdt_wide$d_both_avg, paired = TRUE)
diff_d  <- mean(sdt_wide$d_single_avg - sdt_wide$d_both_avg)
dz_d    <- diff_d / sd(sdt_wide$d_single_avg - sdt_wide$d_both_avg)

cat(sprintf("Main effect of OtherTypePresent on d':\n"))
cat(sprintf("  Single: M = %.3f, Both: M = %.3f\n",
            mean(sdt_wide$d_single_avg), mean(sdt_wide$d_both_avg)))
cat(sprintf("  Diff (single - both) = %+.3f\n", diff_d))
cat(sprintf("  t(%d) = %.3f, p = %.4f, dz = %.3f\n\n",
            t_otp_d$parameter, t_otp_d$statistic, t_otp_d$p.value, dz_d))

# Main effect of ErrorType on d'
sdt_wide <- sdt_wide %>%
  mutate(
    d_phon_avg   = (d_prime_Phonological_Single + d_prime_Phonological_Both) / 2,
    d_nonphon_avg = (`d_prime_Non-Phonological_Single` + `d_prime_Non-Phonological_Both`) / 2
  )

t_et_d <- t.test(sdt_wide$d_phon_avg, sdt_wide$d_nonphon_avg, paired = TRUE)
cat(sprintf("Main effect of ErrorType on d':\n"))
cat(sprintf("  Phon: M = %.3f, NonPhon: M = %.3f\n",
            mean(sdt_wide$d_phon_avg), mean(sdt_wide$d_nonphon_avg)))
cat(sprintf("  t(%d) = %.3f, p = %.4f\n\n",
            t_et_d$parameter, t_et_d$statistic, t_et_d$p.value))

# Interaction: ErrorType x OTP on d'
sdt_wide <- sdt_wide %>%
  mutate(
    otp_effect_d_phon   = d_prime_Phonological_Single - d_prime_Phonological_Both,
    otp_effect_d_nonphon = `d_prime_Non-Phonological_Single` - `d_prime_Non-Phonological_Both`
  )

t_int_d <- t.test(sdt_wide$otp_effect_d_phon, sdt_wide$otp_effect_d_nonphon, paired = TRUE)
cat(sprintf("Interaction (ErrorType x OTP) on d':\n"))
cat(sprintf("  OTP effect on Phon d':    %+.3f\n", mean(sdt_wide$otp_effect_d_phon)))
cat(sprintf("  OTP effect on NonPhon d': %+.3f\n", mean(sdt_wide$otp_effect_d_nonphon)))
cat(sprintf("  t(%d) = %.3f, p = %.4f\n\n",
            t_int_d$parameter, t_int_d$statistic, t_int_d$p.value))

# Simple effects of OTP within each error type
cat("Simple effects of OTP on d':\n")
t_phon <- t.test(sdt_wide$d_prime_Phonological_Single,
                  sdt_wide$d_prime_Phonological_Both, paired = TRUE)
cat(sprintf("  Phon: single M=%.3f, both M=%.3f, diff=%+.3f, t=%.3f, p=%.4f\n",
            mean(sdt_wide$d_prime_Phonological_Single),
            mean(sdt_wide$d_prime_Phonological_Both),
            mean(sdt_wide$d_prime_Phonological_Single - sdt_wide$d_prime_Phonological_Both),
            t_phon$statistic, t_phon$p.value))

t_np <- t.test(sdt_wide$`d_prime_Non-Phonological_Single`,
                sdt_wide$`d_prime_Non-Phonological_Both`, paired = TRUE)
cat(sprintf("  NonPhon: single M=%.3f, both M=%.3f, diff=%+.3f, t=%.3f, p=%.4f\n\n",
            mean(sdt_wide$`d_prime_Non-Phonological_Single`),
            mean(sdt_wide$`d_prime_Non-Phonological_Both`),
            mean(sdt_wide$`d_prime_Non-Phonological_Single` - sdt_wide$`d_prime_Non-Phonological_Both`),
            t_np$statistic, t_np$p.value))

# CRITERION (c) TESTS
cat("CRITERION (c) TESTS\n\n")

t_otp_c <- t.test(sdt_wide$c_single_avg, sdt_wide$c_both_avg, paired = TRUE)
diff_c  <- mean(sdt_wide$c_both_avg - sdt_wide$c_single_avg)
cat(sprintf("Main effect of OtherTypePresent on c:\n"))
cat(sprintf("  Single: M = %.3f, Both: M = %.3f\n",
            mean(sdt_wide$c_single_avg), mean(sdt_wide$c_both_avg)))
cat(sprintf("  Diff (both - single) = %+.3f\n", diff_c))
cat(sprintf("  t(%d) = %.3f, p = %.4f\n\n",
            t_otp_c$parameter, t_otp_c$statistic, t_otp_c$p.value))

# ErrorType effect on c
sdt_wide <- sdt_wide %>%
  mutate(
    c_phon_avg   = (c_bias_Phonological_Single + c_bias_Phonological_Both) / 2,
    c_nonphon_avg = (`c_bias_Non-Phonological_Single` + `c_bias_Non-Phonological_Both`) / 2
  )

t_et_c <- t.test(sdt_wide$c_phon_avg, sdt_wide$c_nonphon_avg, paired = TRUE)
cat(sprintf("Main effect of ErrorType on c:\n"))
cat(sprintf("  Phon: M = %.3f, NonPhon: M = %.3f\n",
            mean(sdt_wide$c_phon_avg), mean(sdt_wide$c_nonphon_avg)))
cat(sprintf("  t(%d) = %.3f, p = %.4f\n\n",
            t_et_c$parameter, t_et_c$statistic, t_et_c$p.value))

# Interaction on c
sdt_wide <- sdt_wide %>%
  mutate(
    otp_effect_c_phon   = c_bias_Phonological_Both - c_bias_Phonological_Single,
    otp_effect_c_nonphon = `c_bias_Non-Phonological_Both` - `c_bias_Non-Phonological_Single`
  )

t_int_c <- t.test(sdt_wide$otp_effect_c_phon, sdt_wide$otp_effect_c_nonphon, paired = TRUE)
cat(sprintf("Interaction (ErrorType x OTP) on c:\n"))
cat(sprintf("  OTP effect on Phon c:    %+.3f\n", mean(sdt_wide$otp_effect_c_phon)))
cat(sprintf("  OTP effect on NonPhon c: %+.3f\n", mean(sdt_wide$otp_effect_c_nonphon)))
cat(sprintf("  t(%d) = %.3f, p = %.4f\n\n",
            t_int_c$parameter, t_int_c$statistic, t_int_c$p.value))

# Simple effects on c
cat("Simple effects of OTP on c:\n")
t_c_phon <- t.test(sdt_wide$c_bias_Phonological_Single,
                    sdt_wide$c_bias_Phonological_Both, paired = TRUE)
cat(sprintf("  Phon: single M=%.3f, both M=%.3f, diff=%+.3f, t=%.3f, p=%.4f\n",
            mean(sdt_wide$c_bias_Phonological_Single),
            mean(sdt_wide$c_bias_Phonological_Both),
            mean(sdt_wide$c_bias_Phonological_Both - sdt_wide$c_bias_Phonological_Single),
            t_c_phon$statistic, t_c_phon$p.value))

t_c_np <- t.test(sdt_wide$`c_bias_Non-Phonological_Single`,
                  sdt_wide$`c_bias_Non-Phonological_Both`, paired = TRUE)
cat(sprintf("  NonPhon: single M=%.3f, both M=%.3f, diff=%+.3f, t=%.3f, p=%.4f\n",
            mean(sdt_wide$`c_bias_Non-Phonological_Single`),
            mean(sdt_wide$`c_bias_Non-Phonological_Both`),
            mean(sdt_wide$`c_bias_Non-Phonological_Both` - sdt_wide$`c_bias_Non-Phonological_Single`),
            t_c_np$statistic, t_c_np$p.value))

# Interpretation summary
cat("\nINTERPRETATION:\n")
cat("  Resource-competition predicts: d' drops with OTP, c unchanged\n")
cat("  Criterion-shift predicts:      c shifts with OTP, d' unchanged\n")
cat(sprintf("  d' main effect of OTP: p = %.4f (%s)\n",
            t_otp_d$p.value, if (t_otp_d$p.value < .05) "SIGNIFICANT" else "n.s."))
cat(sprintf("  c  main effect of OTP: p = %.4f (%s)\n",
            t_otp_c$p.value, if (t_otp_c$p.value < .05) "SIGNIFICANT" else "n.s."))
```

```{r sdt-decomp-diagnostics}
# Floor/ceiling diagnostics (6 trials per cell)
cat("Floor/ceiling rates (before log-linear correction):\n")
sdt_decomp %>%
  group_by(ErrorType, Context) %>%
  summarise(
    HR_0  = sum(hit_rate_raw == 0, na.rm = TRUE),
    HR_1  = sum(hit_rate_raw == 1, na.rm = TRUE),
    FAR_0 = sum(fa_rate_raw == 0, na.rm = TRUE),
    FAR_1 = sum(fa_rate_raw == 1, na.rm = TRUE),
    N     = n(),
    .groups = "drop"
  ) %>%
  knitr::kable(caption = "Floor/ceiling counts per cell")
```


# 4. Figures

## Figure 1: Sensitivity (d') by ErrorType

```{r figure-1-dprime, fig.width=12, fig.height=6}
# d' plot comparing Phon vs NonPhon across attempts
p_att1_dprime <- extract_pval(m1_att1, "ErrorType1")
p_comb_dprime <- extract_pval(m1_comb, "ErrorType1")

stat_dprime <- tibble(
  Attempt    = c("Attempt 1", "Combined (Attempts 1 & 2)"),
  group1     = "Phonological",
  group2     = "Non-Phonological",
  p_label    = c(fmt_p(p_att1_dprime), fmt_p(p_comb_dprime)),
  y.position = max(master_sdt$d_prime, na.rm = TRUE) + 0.3
)

p1 <- ggplot(master_sdt, aes(x = ErrorType, y = d_prime, fill = ErrorType)) +
  geom_boxplot(width = 0.55, outlier.shape = NA, alpha = 0.75,
               color = "grey30", linewidth = 0.55) +
  geom_jitter(width = 0.12, size = 2.0, alpha = 0.40,
              color = "grey25", show.legend = FALSE) +
  stat_pvalue_manual(stat_dprime, label = "p_label", y.position = "y.position",
                     tip.length = 0.015, size = 5.5, bracket.size = 0.7,
                     inherit.aes = FALSE) +
  facet_wrap(~ Attempt, labeller = labeller(Attempt = c(
    "Attempt 1"                 = "Attempt 1 (Initial)",
    "Combined (Attempts 1 & 2)" = "Combined (After Re-exposure)"
  ))) +
  scale_fill_manual(values = error_colors) +
  scale_x_discrete(labels = c(
    "Phonological" = "Phonological",
    "Non-Phonological" = "Non-\nPhonological"
  )) +
  labs(title = "Sensitivity (d')", y = "d'", x = NULL, fill = "Error Type",
       caption = "p-value: ErrorType main effect from GLMER | d' = 0: chance") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)))

p1
ggsave("figure1_sensitivity.tiff", p1, width = 12, height = 6, dpi = 600)
```

## Figure 2: Masking / Competition

```{r figure-2-masking, fig.width=12, fig.height=6}
# Build masking-specific data: detection rate by ErrorType x Context
build_masking_df <- function(df_long) {
  df_long %>%
    filter(Present == 1) %>%
    mutate(
      Context = if_else(OtherTypePresent == 0,
                        "Single Error Type", "Both Error Types")
    ) %>%
    group_by(participant_id, ErrorType, Context) %>%
    summarise(hit_rate = mean(Reported), .groups = "drop")
}

calc_masking_summary <- function(mask_df) {
  mask_df %>%
    group_by(ErrorType, Context) %>%
    summarise(
      stats = list(smean.cl.boot(hit_rate)),
      .groups = "drop"
    ) %>%
    unnest_wider(stats)
}

# First attempt
master_mask_att1 <- build_masking_df(df_long_att1) %>%
  mutate(Attempt = "Attempt 1")
mask_means_att1  <- calc_masking_summary(master_mask_att1) %>%
  mutate(Attempt = "Attempt 1")

# Combined
master_mask_comb <- build_masking_df(df_long_comb) %>%
  mutate(Attempt = "Combined (Attempts 1 & 2)")
mask_means_comb  <- calc_masking_summary(master_mask_comb) %>%
  mutate(Attempt = "Combined (Attempts 1 & 2)")

master_mask <- bind_rows(master_mask_att1, master_mask_comb)
mask_means  <- bind_rows(mask_means_att1, mask_means_comb)

# Masking p-values from GLMER (OtherTypePresent_c effect)
p_mask_att1 <- extract_pval(m_glmer, "OtherTypePresent_c")
p_mask_comb <- {
  m_comb <- glmer(
    Reported ~ ErrorType1 * Present_c + ErrorType1 * OtherTypePresent_c +
      (1 | participant_id) + (1 | question),
    data    = df_long_comb,
    family  = binomial,
    control = glmerControl(optimizer = "bobyqa")
  )
  extract_pval(m_comb, "OtherTypePresent_c")
}

stat_mask <- tibble(
  Attempt    = c("Attempt 1", "Combined (Attempts 1 & 2)"),
  group1     = "Single Error Type",
  group2     = "Both Error Types",
  p_label    = c(fmt_p(p_mask_att1), fmt_p(p_mask_comb)),
  y.position = c(1.05, 1.05)
)

p2 <- ggplot(mask_means, aes(x = Context, y = Mean,
                              color = ErrorType, group = ErrorType)) +
  geom_jitter(data = master_mask, aes(y = hit_rate),
              width = 0.06, height = 0, size = 1.8, alpha = 0.25,
              show.legend = FALSE) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper, fill = ErrorType),
              alpha = 0.18, color = NA) +
  geom_line(linewidth = 1.8) +
  geom_point(size = 5, shape = 19) +
  stat_pvalue_manual(stat_mask, label = "p_label", y.position = "y.position",
                     tip.length = 0.015, size = 8, bracket.size = 0.7,
                     inherit.aes = FALSE) +
  facet_wrap(~ Attempt, labeller = labeller(Attempt = c(
    "Attempt 1"                 = "Attempt 1 (Initial)",
    "Combined (Attempts 1 & 2)" = "Combined (After Re-exposure)"
  ))) +
  scale_color_manual(values = error_colors) +
  scale_fill_manual(values = error_colors) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1.15),
                     breaks = seq(0, 1, 0.25)) +
  scale_x_discrete(labels = c("Single Error Type" = "Single\nError Type",
                               "Both Error Types"  = "Both\nError Types")) +
  labs(y = "Detection Rate", x = NULL, color = "Error Type", fill = "Error Type",
       caption = "Downward slope = masking (detection cost of adding a second error)") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 5)),
         fill = "none") +
  theme(legend.text = element_text(size = 14), legend.title = element_text(size = 14))

p2
ggsave("figure2_masking.tiff", p2, width = 12, height = 6, dpi = 600)
```

```{r fig1-fig2-combined, fig.width=18, fig.height=8}
p3 <- (p1 | p2) +
  plot_annotation(
    tag_levels = "A",
    theme = theme(plot.tag = element_text(size = 24, face = "bold"))
  )
p3
```

## Figure 3: SDT Competition Decomposition (d' and c by ErrorType x Context)

```{r figure-3-sdt-decomposition, fig.width=14, fig.height=6}
# Panel A: d'
decomp_d_summary <- sdt_decomp %>%
  group_by(ErrorType, Context) %>%
  summarise(
    M  = mean(d_prime),
    SE = sd(d_prime) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    ErrorType = factor(ErrorType, levels = c("Phonological", "Non-Phonological")),
    Context   = factor(Context, levels = c("Single", "Both"))
  )

pa <- ggplot(decomp_d_summary,
             aes(x = Context, y = M, fill = ErrorType, alpha = Context)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7,
           color = "white", linewidth = 0.8) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE),
                position = position_dodge(width = 0.8), width = 0.2,
                linewidth = 0.8) +
  scale_fill_manual(values = error_colors) +
  scale_alpha_manual(values = c("Single" = 0.9, "Both" = 0.5), guide = "none") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", linewidth = 0.5) +
  labs(title = "A. Sensitivity (d')", y = "d'", x = "Error Context",
       fill = "Error Type") +
  theme(legend.position = "top")

# Panel B: c
decomp_c_summary <- sdt_decomp %>%
  group_by(ErrorType, Context) %>%
  summarise(
    M  = mean(c_bias),
    SE = sd(c_bias) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    ErrorType = factor(ErrorType, levels = c("Phonological", "Non-Phonological")),
    Context   = factor(Context, levels = c("Single", "Both"))
  )

pb <- ggplot(decomp_c_summary,
             aes(x = Context, y = M, fill = ErrorType, alpha = Context)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7,
           color = "white", linewidth = 0.8) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE),
                position = position_dodge(width = 0.8), width = 0.2,
                linewidth = 0.8) +
  scale_fill_manual(values = error_colors) +
  scale_alpha_manual(values = c("Single" = 0.9, "Both" = 0.5), guide = "none") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", linewidth = 0.5) +
  annotate("text", x = 0.5, y = 0.02, label = "Neutral",
           fontface = "italic", size = 4, color = "grey40") +
  labs(title = "B. Criterion (c)", y = "c (bias)", x = "Error Context",
       fill = "Error Type",
       caption = "Positive c = conservative (reluctant to report)") +
  theme(legend.position = "top")

p_decomp <- pa | pb
p_decomp + plot_annotation(
  title = "SDT Decomposition of Competition Effect",
  subtitle = "Single-error vs. Both-error context",
  theme = theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 13)
  )
)

ggsave("figure3_sdt_decomposition.tiff", p_decomp, width = 14, height = 6, dpi = 600)
```

## Supplementary: Bias (c) by ErrorType

```{r appendix-bias-plot, fig.width=10, fig.height=6}
p_att1_bias <- extract_pval(m1_att1, "ErrorType1")
p_comb_bias <- extract_pval(m1_comb, "ErrorType1")

stat_bias <- tibble(
  Attempt    = c("Attempt 1", "Combined (Attempts 1 & 2)"),
  group1     = "Phonological",
  group2     = "Non-Phonological",
  p_label    = c(fmt_p(p_att1_bias), fmt_p(p_comb_bias)),
  y.position = max(master_sdt$c_bias, na.rm = TRUE) + 0.25
)

p4 <- ggplot(master_sdt, aes(x = ErrorType, y = c_bias, fill = ErrorType)) +
  geom_boxplot(width = 0.55, outlier.shape = NA, alpha = 0.75,
               color = "grey30", linewidth = 0.55) +
  geom_jitter(width = 0.12, size = 2.0, alpha = 0.40,
              color = "grey25", show.legend = FALSE) +
  stat_pvalue_manual(stat_bias, label = "p_label", y.position = "y.position",
                     tip.length = 0.015, size = 5.5, bracket.size = 0.7,
                     inherit.aes = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", linewidth = 0.7) +
  annotate("text", x = 0.55, y = 0, label = "Neutral",
           vjust = -0.6, fontface = "italic", size = 5, color = "grey40") +
  facet_wrap(~ Attempt, labeller = labeller(Attempt = c(
    "Attempt 1"                 = "Attempt 1\n(Initial)",
    "Combined (Attempts 1 & 2)" = "Combined\n(After Re-exposure)"
  ))) +
  scale_fill_manual(values = error_colors) +
  scale_x_discrete(labels = c("Phonological" = "Phonological",
                               "Non-Phonological" = "Non-\nPhonological")) +
  labs(title = "Response Bias (c)", y = "Bias (c)", x = NULL, fill = "Error Type",
       subtitle = "Positive = conservative (reluctant to report) | Negative = liberal",
       caption = "p-value: ErrorType main effect from GLMER | c = 0: neutral bias") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)))

p4
```


# 5. Stopping Rule Analysis

In both_err trials, do participants detect both errors or stop after detecting one?

```{r stopping-rule}
both_err_trials <- df_trials %>% filter(condition == "both_err")

cat(sprintf("Both_err trials: %d from %d participants\n\n",
            nrow(both_err_trials), n_distinct(both_err_trials$participant_id)))

# Classify first-attempt responses
stopping_first <- both_err_trials %>%
  mutate(
    response_pattern = case_when(
      err_type_first %in% c("both", "both?")          ~ "detected_both",
      err_type_first %in% c("p", "p?", "otherp")      ~ "only_phon",
      err_type_first %in% c("np", "np?", "othernp")   ~ "only_nonphon",
      err_reported_first == "y"                        ~ "reported_other",
      TRUE                                             ~ "missed_both"
    )
  )

cat("First-attempt response patterns in both_err condition:\n")
stopping_first %>%
  count(response_pattern) %>%
  mutate(pct = 100 * n / sum(n)) %>%
  arrange(desc(n)) %>%
  knitr::kable(digits = 1)

# Of those who detected at least one, what fraction detected both?
detected_any <- stopping_first %>%
  filter(response_pattern %in% c("detected_both", "only_phon", "only_nonphon"))

if (nrow(detected_any) > 0) {
  n_both <- sum(detected_any$response_pattern == "detected_both")
  n_one  <- sum(detected_any$response_pattern != "detected_both")
  cat(sprintf("\nOf trials with >=1 error detected (N=%d):\n", nrow(detected_any)))
  cat(sprintf("  Detected both: %d (%.1f%%)\n", n_both, 100 * n_both / nrow(detected_any)))
  cat(sprintf("  Stopped at one: %d (%.1f%%)\n", n_one, 100 * n_one / nrow(detected_any)))
  cat(sprintf("    Only phon:    %d\n", sum(detected_any$response_pattern == "only_phon")))
  cat(sprintf("    Only nonphon: %d\n", sum(detected_any$response_pattern == "only_nonphon")))
}

# Combined (first + second attempt)
stopping_comb <- both_err_trials %>%
  mutate(
    detected_phon_comb = phon_detected_comb,
    detected_nonphon_comb = nonphon_detected_comb,
    response_pattern_comb = case_when(
      detected_phon_comb & detected_nonphon_comb ~ "detected_both",
      detected_phon_comb                         ~ "only_phon",
      detected_nonphon_comb                      ~ "only_nonphon",
      err_reported_first == "y" | err_reported_second == "y" ~ "reported_other",
      TRUE                                       ~ "missed_both"
    )
  )

cat("\nCombined (first + second attempt) response patterns:\n")
stopping_comb %>%
  count(response_pattern_comb) %>%
  mutate(pct = 100 * n / sum(n)) %>%
  arrange(desc(n)) %>%
  knitr::kable(digits = 1)
```


# 6. Learning Effects in Error Detection

Does detection rate improve across the 24 target trials?

```{r learning-detection}
# Trial position effects on detection
learning_df <- df_long_att1 %>%
  filter(Present == 1) %>%
  left_join(
    df_trials %>% select(participant_id, question, trial_position),
    by = c("participant_id", "question")
  )

# Overall trend
m_learn <- glmer(
  Reported ~ scale(trial_position) + ErrorType1 +
    (1 | participant_id) + (1 | question),
  data    = learning_df,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

cat("Learning effect in error detection:\n")
cat("GLMER: Reported ~ trial_position + ErrorType + (1|participant) + (1|question)\n\n")
summary(m_learn)

# Mean detection rate by trial quartile
learning_df %>%
  mutate(quartile = cut(trial_position, breaks = c(0, 6, 12, 18, 24),
                        labels = c("Q1 (1-6)", "Q2 (7-12)", "Q3 (13-18)", "Q4 (19-24)"))) %>%
  group_by(quartile, ErrorType) %>%
  summarise(detection_rate = mean(Reported), n = n(), .groups = "drop") %>%
  pivot_wider(names_from = ErrorType, values_from = c(detection_rate, n)) %>%
  knitr::kable(digits = 3, caption = "Detection rate by trial position quartile")
```


# 7. Sequential Trial Effects

Does the condition or detection outcome on trial N affect trial N+1?

```{r sequential-effects}
# Build transition data from consecutive target trials
transitions <- df_trials %>%
  arrange(participant_id, trial_position) %>%
  group_by(participant_id) %>%
  mutate(
    # Previous trial info
    prev_condition     = lag(condition),
    prev_reported      = lag(err_reported_first == "y"),
    prev_type          = lag(err_type_first),
    prev_phon_detected = lag(phon_detected_first),
    prev_nonphon_detected = lag(nonphon_detected_first)
  ) %>%
  ungroup() %>%
  filter(!is.na(prev_condition))

cat(sprintf("Total target-to-target transitions: %d from %d participants\n\n",
            nrow(transitions), n_distinct(transitions$participant_id)))

# Analysis 1: Does reporting on trial N predict reporting on N+1?
any_detected_curr <- transitions %>%
  mutate(reported_any = err_reported_first == "y")

after_report <- any_detected_curr %>%
  filter(prev_reported == TRUE) %>%
  summarise(rate = mean(reported_any), n = n())

after_no_report <- any_detected_curr %>%
  filter(prev_reported == FALSE) %>%
  summarise(rate = mean(reported_any), n = n())

# Chi-squared test
tab <- table(
  Previous = any_detected_curr$prev_reported,
  Current  = any_detected_curr$reported_any
)
chi_res <- chisq.test(tab)

cat("Does reporting on trial N predict reporting on N+1?\n")
cat(sprintf("  After REPORTING error:     %.3f report on next (N=%d)\n",
            after_report$rate, after_report$n))
cat(sprintf("  After NOT reporting error: %.3f report on next (N=%d)\n",
            after_no_report$rate, after_no_report$n))
cat(sprintf("  Chi-squared: X2=%.3f, p=%.4f\n\n", chi_res$statistic, chi_res$p.value))

# Analysis 2: Does error TYPE on trial N affect what's reported on N+1?
cat("Does error TYPE on trial N affect detection on N+1?\n")

prev_type_effects <- any_detected_curr %>%
  mutate(
    prev_category = case_when(
      prev_reported & prev_type %in% c("p", "p?")                  ~ "Phon detected",
      prev_reported & prev_type %in% c("np", "np?")                ~ "NonPhon detected",
      prev_reported & prev_type %in% c("both", "both?")            ~ "Both detected",
      !prev_reported                                                ~ "Nothing detected",
      TRUE                                                          ~ "Other"
    )
  ) %>%
  filter(prev_category != "Other") %>%
  group_by(prev_category) %>%
  summarise(
    next_report_rate = mean(reported_any),
    n = n(),
    .groups = "drop"
  )

prev_type_effects %>%
  knitr::kable(digits = 3, caption = "Next-trial report rate by previous detection type")
```

```{r randomization-check}
# Was trial order randomized?
item_positions <- df_trials %>%
  group_by(question) %>%
  summarise(
    n_participants = n(),
    n_positions    = n_distinct(trial_position),
    min_pos = min(trial_position),
    max_pos = max(trial_position),
    .groups = "drop"
  ) %>%
  filter(n_participants >= 5)

n_varied <- sum(item_positions$n_positions > 1)
cat(sprintf("\nRandomization check:\n"))
cat(sprintf("  Items with >=5 participants: %d\n", nrow(item_positions)))
cat(sprintf("  Items appearing in different positions: %d\n", n_varied))
cat(sprintf("  -> %s\n",
            if (n_varied / nrow(item_positions) > 0.5)
              "YES, trial order was randomized"
            else "Trial order appears FIXED"))
```


# 8. Re-exposure / Replay Analysis

```{r reexposure-rates}
# Count re-exposure usage per participant
reexposure_summary <- df_trials %>%
  group_by(participant_id) %>%
  summarise(
    total_trials = n(),
    n_replayed   = sum(second_attempt_used),
    replay_rate  = n_replayed / total_trials,
    .groups = "drop"
  )

cat("Re-exposure usage:\n")
cat(sprintf("  Participants: %d\n", nrow(reexposure_summary)))
cat(sprintf("  Replay rate: M=%.3f, SD=%.3f, median=%.3f, range=[%.3f, %.3f]\n",
            mean(reexposure_summary$replay_rate),
            sd(reexposure_summary$replay_rate),
            median(reexposure_summary$replay_rate),
            min(reexposure_summary$replay_rate),
            max(reexposure_summary$replay_rate)))
cat(sprintf("  Total replays per participant: M=%.1f, SD=%.1f, range=[%d, %d]\n",
            mean(reexposure_summary$n_replayed),
            sd(reexposure_summary$n_replayed),
            min(reexposure_summary$n_replayed),
            max(reexposure_summary$n_replayed)))
```

```{r reexposure-by-condition}
# Replay rates by condition
reexposure_by_cond <- df_trials %>%
  group_by(condition) %>%
  summarise(
    n_trials   = n(),
    n_replayed = sum(second_attempt_used),
    replay_rate = n_replayed / n_trials,
    .groups = "drop"
  )

cat("\nReplay rate by condition:\n")
reexposure_by_cond %>%
  knitr::kable(digits = 3)
```

```{r reexposure-detection-change}
# Does re-exposure improve detection?
# Compare first-attempt vs. combined detection rates for participants who replayed

replay_effect <- df_trials %>%
  filter(second_attempt_used, condition != "correct") %>%
  mutate(
    detected_first = case_when(
      condition == "phon_err"     ~ phon_detected_first,
      condition == "non_phon_err" ~ nonphon_detected_first,
      condition == "both_err"     ~ phon_detected_first | nonphon_detected_first
    ),
    detected_comb = case_when(
      condition == "phon_err"     ~ phon_detected_comb,
      condition == "non_phon_err" ~ nonphon_detected_comb,
      condition == "both_err"     ~ phon_detected_comb | nonphon_detected_comb
    ),
    gained_on_replay = !detected_first & detected_comb
  )

if (nrow(replay_effect) > 0) {
  cat("\nRe-exposure effect on detection (trials where replay was used):\n")
  cat(sprintf("  Trials with replay in error conditions: %d\n", nrow(replay_effect)))
  cat(sprintf("  First-attempt detection rate: %.3f\n", mean(replay_effect$detected_first)))
  cat(sprintf("  Combined detection rate:      %.3f\n", mean(replay_effect$detected_comb)))
  cat(sprintf("  Gained on replay:             %.3f (%d trials)\n",
              mean(replay_effect$gained_on_replay), sum(replay_effect$gained_on_replay)))
}

# ErrorType x Presentation (first vs combined) interaction
# Are non-phonological errors more likely to be caught on replay?
replay_by_type <- df_long_att1 %>%
  filter(Present == 1) %>%
  select(participant_id, question, ErrorType, Reported_first = Reported) %>%
  left_join(
    df_long_comb %>%
      filter(Present == 1) %>%
      select(participant_id, question, ErrorType, Reported_comb = Reported),
    by = c("participant_id", "question", "ErrorType")
  ) %>%
  pivot_longer(
    cols = c(Reported_first, Reported_comb),
    names_to = "Presentation",
    values_to = "Reported"
  ) %>%
  mutate(
    Presentation = if_else(Presentation == "Reported_first",
                           "First", "Combined"),
    Presentation_c = if_else(Presentation == "Combined", 0.5, -0.5),
    ErrorType1 = if_else(ErrorType == "Non-Phonological", 1, -1)
  )

m_replay <- glmer(
  Reported ~ ErrorType1 * Presentation_c + (1 | participant_id) + (1 | question),
  data    = replay_by_type,
  family  = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

cat("\nErrorType x Presentation interaction (first vs combined):\n")
summary(m_replay)

p_interaction <- extract_pval(m_replay, "ErrorType1:Presentation_c")
cat(sprintf("\n  ErrorType x Presentation interaction: p = %.4f\n", p_interaction))
```


# 9. Transcription Analysis

```{r transcription-helpers}
# Text substitution helper
mgsub <- function(pattern, replacement, x, ...) {
  result <- x
  for (i in seq_along(pattern))
    result <- gsub(pattern[i], replacement[i], result, ...)
  result
}

# Keyword scoring (pmatch-based)
num.keywords.correct <- function(transcription, keywords) {
  keys  <- gsub("[[:punct:]]", "", tolower(as.character(keywords)))
  trans <- gsub("[[:punct:]]", "", tolower(as.character(transcription)))
  sum(!is.na(pmatch(unlist(strsplit(keys,  " ")),
                    unlist(strsplit(trans, " ")))))
}

# Per-keyword scoring for detailed analysis
check_keywords_correct <- function(transcription, keywords,
                                    global_row_id, filename, participant_id) {
  keys  <- gsub("[[:punct:]]", "", tolower(as.character(keywords)))
  trans <- gsub("[[:punct:]]", "", tolower(as.character(transcription)))
  keys.list  <- unlist(strsplit(keys,  " "))
  trans.list <- unlist(strsplit(trans, " "))
  data.frame(
    global_row_id  = global_row_id,
    participant_id = participant_id,
    Filename       = filename,
    keyword        = keys.list,
    is_correct     = keys.list %in% trans.list
  )
}
```

```{r transcription-load}
# Load transcription data
process_transcription_batch <- function(folder_name, group_label,
                                         batch_label, has_questionnaire) {
  target_folder <- file.path("../data/preprocessed", folder_name)
  trans_files <- list.files(target_folder, pattern = "transcription_data.csv",
                             full.names = TRUE, recursive = TRUE)
  if (length(trans_files) == 0) { warning(paste("! No data for", batch_label)); return(NULL) }

  d_combined <- map_dfr(trans_files, function(f_path) {
    read_csv(f_path, show_col_types = FALSE) %>% mutate(across(everything(), as.character))
  })

  key_file <- "../data/transcription_scoring.txt"
  if (!file.exists(key_file)) { warning("Key file missing"); return(NULL) }
  key.df <- read.delim(key_file, header = TRUE, sep = ",") %>%
    mutate(Sentence = as.character(Sentence))

  d.trans <- d_combined %>%
    filter(!is.na(stimulus_files_presented), stimulus_files_presented != "",
           block == "Transcription") %>%
    mutate(Filename      = basename(as.character(stimulus_files_presented)),
           Transcription = response_value)

  d.all.m <- merge(d.trans, subset(key.df, select = -c(SentenceSet)),
                   by.x = "Filename", by.y = "Sentence", sort = FALSE)
  if (nrow(d.all.m) == 0) { warning(paste("! Merge failed for", batch_label)); return(NULL) }

  d.all.m <- within(d.all.m, {
    Transcription <- tolower(mgsub(
      c("&lt&","&gt&","&quot&","&comma&","&semicol&","&pipe&","&apos&"),
      c("<",   ">",   '"',     ",",      ";",         "|",     "'"),
      Transcription))
    NumKeywordsCorrect  <- mapply(num.keywords.correct, Transcription, Keywords)
    PropKeywordsCorrect <- NumKeywordsCorrect / as.numeric(NumKeywords)
  })

  # Add questionnaire data
  d.all.m$quest_exposure_spanish_accent <- NA_character_
  if (has_questionnaire) {
    q_files <- list.files(target_folder, pattern = "questionnaire_data.csv",
                           full.names = TRUE, recursive = TRUE)
    if (length(q_files) > 0) {
      df_q <- map_dfr(q_files, function(f) {
        read_csv(f, show_col_types = FALSE) %>%
          filter(stimuli_presented %in% c("exposure_spanish_accent", "exposure_spanish")) %>%
          select(participant_id, question_id = stimuli_presented, answer = response_value) %>%
          distinct() %>%
          pivot_wider(names_from = question_id, values_from = answer) %>%
          rename_with(~ paste0("quest_", .), .cols = -participant_id)
      }) %>% group_by(participant_id) %>% slice(1) %>% ungroup()

      d.all.m <- left_join(d.all.m, df_q, by = "participant_id")
      if ("quest_exposure_spanish_accent.y" %in% names(d.all.m)) {
        d.all.m$quest_exposure_spanish_accent <- d.all.m$quest_exposure_spanish_accent.y
        d.all.m$quest_exposure_spanish_accent.x <- NULL
        d.all.m$quest_exposure_spanish_accent.y <- NULL
      }
    }
  }

  d.all.m$batch_id <- batch_label
  d.all.m$group    <- group_label
  d.all.m
}

# Load both batches
batch_manifest_trans <- tribble(
  ~folder_name,         ~group_label, ~batch_label,   ~has_questionnaire,
  "transcrip_baseline", "baseline",   "Baseline",     TRUE,
  "main_batch3",        "main",       "Main Batch 3", TRUE
)

df_transcription_raw <- pmap_dfr(batch_manifest_trans, process_transcription_batch)

# Filter to eligible participants (exclude attention check failures)
eligible_ids <- unique(df_trials$participant_id)

df_transcription_all <- df_transcription_raw %>%
  filter(batch_id == "Baseline" | participant_id %in% eligible_ids) %>%
  # Also exclude one problematic participant from baseline
  filter(participant_id !=
           "6df14d8680851e21a1bbe20d8e32fc04bc30979c5d654dc35ef36bc2910b9765") %>%
  group_by(participant_id) %>%
  mutate(trial_order = row_number()) %>%
  ungroup()

df_transcription_all %>%
  group_by(batch_id) %>%
  summarise(N = n_distinct(participant_id), Trials = n()) %>%
  knitr::kable(caption = "Transcription data summary")
```

## 9.1 Group Comparison (Baseline vs. Error-Detection)

```{r transcription-group-comparison, fig.width=10, fig.height=6}
# Per-participant means
pts <- df_transcription_all %>%
  group_by(batch_id, participant_id) %>%
  summarise(M = mean(PropKeywordsCorrect, na.rm = TRUE), .groups = "drop")

# Bootstrapped group means
summ_group <- df_transcription_all %>%
  group_by(batch_id) %>%
  summarise(stats = list(smean.cl.boot(PropKeywordsCorrect)), .groups = "drop") %>%
  unnest_wider(stats)

p5 <- ggplot(summ_group, aes(x = batch_id, y = Mean, color = batch_id)) +
  geom_jitter(data = pts, aes(y = M),
              width = 0.12, alpha = 0.25, size = 2, color = "grey55") +
  geom_errorbar(aes(ymin = Lower, ymax = Upper),
                width = 0.12, linewidth = 1.3) +
  geom_point(size = 6, shape = 19) +
  geom_text(aes(label = percent(Mean, accuracy = 0.1)),
            hjust = -0.45, fontface = "bold", size = 5.5) +
  scale_color_manual(values = c("Baseline" = "#777777", "Main Batch 3" = COL_NONPHON)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(title = "Accent Adaptation: Group Comparison",
       subtitle = "Mean keyword accuracy | Error bars = 95% bootstrap CI",
       y = "Mean Accuracy", x = NULL) +
  theme(legend.position = "none")

p5
```

```{r transcription-stats}
# Welch t-test (participant-level means)
df_agg <- df_transcription_all %>%
  group_by(participant_id, batch_id) %>%
  summarise(Mean_Accuracy = mean(PropKeywordsCorrect, na.rm = TRUE), .groups = "drop") %>%
  mutate(batch_id = factor(batch_id, levels = c("Baseline", "Main Batch 3")))

t_res <- t.test(Mean_Accuracy ~ batch_id, data = df_agg, var.equal = FALSE)
eff   <- df_agg %>% cohens_d(Mean_Accuracy ~ batch_id, var.equal = FALSE)

cat("Welch t-test (participant-level means):\n")
print(t_res)
cat(sprintf("Cohen's d: %.3f\n\n", eff$effsize))

# LMM: Accuracy ~ Group + (1|Participant) + (1|Sentence)
df_lmm <- df_transcription_all %>%
  mutate(
    batch_id       = factor(batch_id, levels = c("Baseline", "Main Batch 3")),
    participant_id = factor(participant_id),
    Filename       = factor(Filename)
  )

lmm_trans <- lmer(
  PropKeywordsCorrect ~ batch_id + (1 | participant_id) + (1 | Filename),
  data = df_lmm
)

cat("LMM: Accuracy ~ Group + (1|Participant) + (1|Sentence)\n")
summary(lmm_trans)

coefs <- fixef(lmm_trans)
slope <- coefs["batch_idMain Batch 3"]
p_lmm <- summary(lmm_trans)$coefficients["batch_idMain Batch 3", "Pr(>|t|)"]
cat(sprintf("\nBaseline accuracy: %.1f%%\n", 100 * coefs["(Intercept)"]))
cat(sprintf("Adaptation effect: %+.1f%%\n", 100 * slope))
cat(sprintf("LMM p-value: %.4f (%s)\n", p_lmm, if (p_lmm < .05) "SIGNIFICANT" else "n.s."))
```

## 9.2 Learning Effects in Transcription

```{r transcription-learning, fig.width=12, fig.height=6}
# Does transcription accuracy improve across the 20 sentences?
learning_trans <- df_transcription_all %>%
  group_by(batch_id, trial_order) %>%
  summarise(
    M  = mean(PropKeywordsCorrect, na.rm = TRUE),
    SE = sd(PropKeywordsCorrect, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Per-group linear trends
for (grp in unique(df_transcription_all$batch_id)) {
  grp_data <- df_transcription_all %>% filter(batch_id == grp)
  lm_fit <- lm(PropKeywordsCorrect ~ trial_order, data = grp_data)
  slope  <- coef(lm_fit)["trial_order"]
  p_val  <- summary(lm_fit)$coefficients["trial_order", "Pr(>|t|)"]
  r_val  <- sqrt(summary(lm_fit)$r.squared)
  cat(sprintf("%s: slope=%.4f, r=%.3f, p=%.4f (accuracy %s by %.2f pp per sentence)\n",
              grp, slope, r_val, p_val,
              if (slope > 0) "improves" else "declines",
              abs(slope) * 100))
}

p_learn <- ggplot(learning_trans, aes(x = trial_order, y = M, color = batch_id)) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.3, alpha = 0.5) +
  geom_point(size = 3) +
  geom_line(linewidth = 1) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", linewidth = 0.8) +
  scale_color_manual(values = c("Baseline" = "#777777", "Main Batch 3" = COL_NONPHON)) +
  scale_y_continuous(labels = percent_format(), limits = c(0.3, 0.9)) +
  scale_x_continuous(breaks = 1:20) +
  labs(title = "Learning Effect During Transcription",
       x = "Sentence Position (1-20)", y = "Proportion Keywords Correct",
       color = "Group") +
  theme(legend.position = "top")

p_learn
ggsave("figure_learning_effect.tiff", p_learn, width = 12, height = 6, dpi = 300)
```

## 9.3 Hardest Keywords

```{r hardest-keywords, fig.width=12, fig.height=8}
# Keyword-level analysis
df_kw <- df_transcription_all %>%
  mutate(global_row_id = row_number())

kw_df <- pmap_dfr(list(
  df_kw$Transcription, df_kw$Keywords,
  df_kw$global_row_id, df_kw$Filename, df_kw$participant_id
), check_keywords_correct) %>%
  left_join(df_kw %>% select(global_row_id, batch_id), by = "global_row_id")

hardest <- kw_df %>%
  group_by(keyword, batch_id) %>%
  summarise(P = mean(is_correct, na.rm = TRUE), .groups = "drop") %>%
  group_by(batch_id) %>%
  arrange(P) %>%
  slice_head(n = 10)

p6 <- ggplot(hardest,
             aes(x = reorder_within(keyword, P, batch_id), y = P, fill = batch_id)) +
  geom_col(color = "white", alpha = 0.9) +
  scale_x_reordered() +
  facet_wrap(~ batch_id, scales = "free_y") +
  coord_flip() +
  scale_fill_manual(values = c("Baseline" = "#888888", "Main Batch 3" = COL_NONPHON)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(title = "Hardest Keywords by Group", x = NULL, y = "Proportion Correct") +
  theme(legend.position = "none")

p6
```

## 9.4 Accent-Relevant vs. Non-Accent Keywords

```{r accent-keyword-group-comparison, fig.width=12, fig.height=6}
# keyword_contrast is defined in Section 11 setup — define it here if not yet available
if (!exists("keyword_contrast")) {
  keyword_contrast <- bind_rows(
    tribble(
      ~keyword,      ~contrast,
      "rip",         "/ɪ/→/i/",
      "ship",        "/ɪ/→/i/",
      "his",         "/ɪ/→/i/",
      "she",         "/ʃ/→/s/",
      "should",      "/ʃ/→/s/",
      "issue",       "/ʃ/→/s/",
      "cloud",       "/k/→/g/",
      "commented",   "/k/→/g/",
      "consume",     "/k/→/g/",
      "colleague",   "/k/→/g/",
      "considered",  "/k/→/g/",
      "closer",      "/k/→/g/",
      "sky",         "/k/→/g/",
      "comes",       "/k/→/g/",
      "scared",      "/k/→/g/",
      "wrath",       "/θ/→/t/",
      "math",        "/θ/→/t/",
      "bath",        "/θ/→/t/",
      "booth",       "/θ/→/t/",
      "something",   "/θ/→/t/",
      "with",        "/θ/→/t/",
      "value",       "/v/→/b/",
      "have",        "/v/→/b/",
      "obvious",     "/v/→/b/",
      "full",        "/ʊ/→/u/",
      "look",        "/ʊ/→/u/"
    ),
    tribble(
      ~keyword,    ~contrast,
      "weep",      "/ɪ/→/i/",
      "feet",      "/ɪ/→/i/",
      "peel",      "/ɪ/→/i/",
      "bean",      "/ɪ/→/i/",
      "lead",      "/ɪ/→/i/",
      "leak",      "/ɪ/→/i/",
      "heat",      "/ɪ/→/i/",
      "greed",     "/ɪ/→/i/",
      "sleep",     "/ɪ/→/i/",
      "cheek",     "/ɪ/→/i/",
      "theme",     "/ɪ/→/i/",
      "heel",      "/ɪ/→/i/",
      "bead",      "/ɪ/→/i/",
      "sloth",     "/ɪ/→/i/"
    )
  )
}

# Tag each keyword as accent-relevant or not
kw_df <- kw_df %>%
  mutate(accent_relevant = keyword %in% keyword_contrast$keyword)

cat(sprintf("Total keyword observations: %d\n", nrow(kw_df)))
cat(sprintf("  Accent-relevant: %d (keywords: %d unique)\n",
            sum(kw_df$accent_relevant),
            n_distinct(kw_df$keyword[kw_df$accent_relevant])))
cat(sprintf("  Non-accent:      %d (keywords: %d unique)\n\n",
            sum(!kw_df$accent_relevant),
            n_distinct(kw_df$keyword[!kw_df$accent_relevant])))

# Per-participant accuracy for accent-relevant keywords only
accent_pp <- kw_df %>%
  filter(accent_relevant) %>%
  group_by(batch_id, participant_id) %>%
  summarise(accuracy = mean(is_correct, na.rm = TRUE), .groups = "drop")

# Per-participant accuracy for non-accent keywords
nonaccent_pp <- kw_df %>%
  filter(!accent_relevant) %>%
  group_by(batch_id, participant_id) %>%
  summarise(accuracy = mean(is_correct, na.rm = TRUE), .groups = "drop")

# Group means
cat("Accent-relevant keywords:\n")
accent_pp %>%
  group_by(batch_id) %>%
  summarise(M = mean(accuracy), SD = sd(accuracy), N = n(), .groups = "drop") %>%
  knitr::kable(digits = 3, caption = "Accent-relevant keyword accuracy by group")

cat("\nNon-accent keywords:\n")
nonaccent_pp %>%
  group_by(batch_id) %>%
  summarise(M = mean(accuracy), SD = sd(accuracy), N = n(), .groups = "drop") %>%
  knitr::kable(digits = 3, caption = "Non-accent keyword accuracy by group")
```

```{r accent-keyword-stats}
# ---- Accent-relevant keywords: group comparison ----
cat("=== Accent-relevant keywords ===\n\n")

t_accent <- t.test(accuracy ~ batch_id, data = accent_pp, var.equal = FALSE)
d_accent <- accent_pp %>%
  mutate(batch_id = factor(batch_id, levels = c("Baseline", "Main Batch 3"))) %>%
  cohens_d(accuracy ~ batch_id, var.equal = FALSE)

cat(sprintf("Welch t-test: t(%.1f) = %.3f, p = %.4f\n",
            t_accent$parameter, t_accent$statistic, t_accent$p.value))
cat(sprintf("Cohen's d = %.3f\n\n", d_accent$effsize))

# LMM on accent-relevant keywords (trial-level)
kw_accent_lmm <- kw_df %>%
  filter(accent_relevant) %>%
  mutate(
    batch_id       = factor(batch_id, levels = c("Baseline", "Main Batch 3")),
    participant_id = factor(participant_id),
    keyword        = factor(keyword)
  )

lmm_accent <- glmer(
  is_correct ~ batch_id + (1 | participant_id) + (1 | keyword),
  data = kw_accent_lmm, family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

cat("GLMER: is_correct ~ batch_id + (1|participant) + (1|keyword)\n")
summary(lmm_accent)

# ---- Non-accent keywords: group comparison ----
cat("\n\n=== Non-accent keywords ===\n\n")

t_nonaccent <- t.test(accuracy ~ batch_id, data = nonaccent_pp, var.equal = FALSE)
d_nonaccent <- nonaccent_pp %>%
  mutate(batch_id = factor(batch_id, levels = c("Baseline", "Main Batch 3"))) %>%
  cohens_d(accuracy ~ batch_id, var.equal = FALSE)

cat(sprintf("Welch t-test: t(%.1f) = %.3f, p = %.4f\n",
            t_nonaccent$parameter, t_nonaccent$statistic, t_nonaccent$p.value))
cat(sprintf("Cohen's d = %.3f\n", d_nonaccent$effsize))
```

```{r accent-keyword-plot, fig.width=12, fig.height=5}
# Side-by-side comparison plot
plot_data <- bind_rows(
  accent_pp %>% mutate(kw_type = "Accent-Relevant"),
  nonaccent_pp %>% mutate(kw_type = "Non-Accent")
)

plot_summ <- plot_data %>%
  group_by(kw_type, batch_id) %>%
  summarise(
    M = mean(accuracy), SE = sd(accuracy) / sqrt(n()),
    .groups = "drop"
  )

p_kwtype <- ggplot(plot_summ, aes(x = batch_id, y = M, fill = batch_id)) +
  geom_col(alpha = 0.8, color = "white", width = 0.6) +
  geom_jitter(data = plot_data, aes(y = accuracy),
              width = 0.12, alpha = 0.2, size = 1.5, color = "grey40") +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.15, linewidth = 0.8) +
  geom_text(aes(label = percent(M, accuracy = 0.1)), vjust = -0.8, fontface = "bold", size = 4.5) +
  facet_wrap(~ kw_type) +
  scale_fill_manual(values = c("Baseline" = "#888888", "Main Batch 3" = COL_NONPHON)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(title = "Transcription Accuracy: Accent-Relevant vs. Non-Accent Keywords",
       subtitle = "Error bars = ±1 SE | Points = individual participants",
       y = "Mean Keyword Accuracy", x = NULL) +
  theme(legend.position = "none")

p_kwtype
```


# 10. Re-exposure x Transcription Interaction

```{r reexposure-transcription}
# Correlate per-participant replay rate with transcription accuracy

trans_means <- df_transcription_all %>%
  filter(batch_id == "Main Batch 3") %>%
  group_by(participant_id) %>%
  summarise(trans_accuracy = mean(PropKeywordsCorrect, na.rm = TRUE), .groups = "drop")

reexp_trans <- reexposure_summary %>%
  inner_join(trans_means, by = "participant_id")

cat(sprintf("Matched participants (replay + transcription): %d\n\n",
            nrow(reexp_trans)))

if (nrow(reexp_trans) >= 10) {
  r_pearson  <- cor.test(reexp_trans$replay_rate, reexp_trans$trans_accuracy)
  r_spearman <- cor.test(reexp_trans$replay_rate, reexp_trans$trans_accuracy,
                          method = "spearman")

  cat("Replay rate x Transcription accuracy:\n")
  cat(sprintf("  Pearson:  r=%.3f, p=%.3f\n", r_pearson$estimate, r_pearson$p.value))
  cat(sprintf("  Spearman: rho=%.3f, p=%.3f\n", r_spearman$estimate, r_spearman$p.value))

  # Total replays x accuracy
  r_total <- cor.test(reexp_trans$n_replayed, reexp_trans$trans_accuracy,
                       method = "spearman")
  cat(sprintf("  Total replay count x accuracy: rho=%.3f, p=%.3f\n",
              r_total$estimate, r_total$p.value))
}
```

```{r reexposure-transcription-plot, fig.width=8, fig.height=6}
if (nrow(reexp_trans) >= 10) {
  p_rxt <- ggplot(reexp_trans, aes(x = replay_rate, y = trans_accuracy)) +
    geom_point(size = 3, alpha = 0.6, color = COL_NONPHON) +
    geom_smooth(method = "lm", se = TRUE, color = "grey40", fill = "grey80") +
    scale_x_continuous(labels = percent_format()) +
    scale_y_continuous(labels = percent_format()) +
    labs(
      title = "Re-exposure Rate vs. Transcription Accuracy",
      subtitle = sprintf("r = %.3f, p = %.3f",
                          r_pearson$estimate, r_pearson$p.value),
      x = "Proportion of Trials Replayed",
      y = "Mean Keyword Accuracy"
    )

  p_rxt
  ggsave("figure_reexposure_x_transcription.tiff", p_rxt,
         width = 8, height = 6, dpi = 300)
}
```


# 11. Similarity Metrics Analysis

```{r sim-setup}
# ---- Accent keyword mappings ----

# Keywords that CHANGE under the Spanish accent (for phonological distance)
changed_keywords <- tribble(
  ~keyword,      ~standard,     ~accented,     ~contrast,
  "rip",         "rip",         "reep",        "/ɪ/→/i/",
  "ship",        "ship",        "sheep",       "/ɪ/→/i/",
  "his",         "his",         "hees",        "/ɪ/→/i/",
  "she",         "she",         "se",          "/ʃ/→/s/",
  "should",      "should",      "sould",       "/ʃ/→/s/",
  "issue",       "issue",       "issoo",       "/ʃ/→/s/",
  "cloud",       "cloud",       "gloud",       "/k/→/g/",
  "commented",   "commented",   "gommented",   "/k/→/g/",
  "consume",     "consume",     "gonsume",     "/k/→/g/",
  "colleague",   "colleague",   "golleague",   "/k/→/g/",
  "considered",  "considered",  "gonsidered",  "/k/→/g/",
  "closer",      "closer",      "gloser",      "/k/→/g/",
  "sky",         "sky",         "sgy",         "/k/→/g/",
  "comes",       "comes",       "gomes",       "/k/→/g/",
  "scared",      "scared",      "sgared",      "/k/→/g/",
  "wrath",       "wrath",       "wrat",        "/θ/→/t/",
  "math",        "math",        "mat",         "/θ/→/t/",
  "bath",        "bath",        "bat",         "/θ/→/t/",
  "booth",       "booth",       "boot",        "/θ/→/t/",
  "something",   "something",   "someting",    "/θ/→/t/",
  "with",        "with",        "wit",         "/θ/→/t/",
  "value",       "value",       "balue",       "/v/→/b/",
  "have",        "have",        "habe",        "/v/→/b/",
  "obvious",     "obvious",     "obbious",     "/v/→/b/",
  "full",        "full",        "fool",        "/ʊ/→/u/",
  "look",        "look",        "luke",        "/ʊ/→/u/"
)

# All accent-relevant keywords → contrast (including unchanged /i:/ words)
keyword_contrast <- bind_rows(
  changed_keywords %>% select(keyword, contrast),
  tribble(
    ~keyword,    ~contrast,
    "weep",      "/ɪ/→/i/",
    "feet",      "/ɪ/→/i/",
    "peel",      "/ɪ/→/i/",
    "bean",      "/ɪ/→/i/",
    "lead",      "/ɪ/→/i/",
    "leak",      "/ɪ/→/i/",
    "heat",      "/ɪ/→/i/",
    "greed",     "/ɪ/→/i/",
    "sleep",     "/ɪ/→/i/",
    "cheek",     "/ɪ/→/i/",
    "theme",     "/ɪ/→/i/",
    "heel",      "/ɪ/→/i/",
    "bead",      "/ɪ/→/i/",
    "sloth",     "/ɪ/→/i/"
  )
)

# Contrast display order and colors
contrast_order <- c("/ɪ/→/i/", "/ʃ/→/s/", "/k/→/g/", "/v/→/b/", "/ʊ/→/u/", "/θ/→/t/")
contrast_colors <- c(
  "/ɪ/→/i/" = "#E07B3F", "/ʃ/→/s/" = "#2B7BB9", "/k/→/g/" = "#57A75C",
  "/v/→/b/" = "#9B6EBD", "/ʊ/→/u/" = "#D4A853", "/θ/→/t/" = "#C75B7A"
)

# Reload scoring key for sentence-level analysis
key.df.sim <- read.delim("../data/transcription_scoring.txt", header = TRUE, sep = ",") %>%
  mutate(Sentence = as.character(Sentence))

# pmatch-style keyword matching helper
is_pmatch <- function(keyword, trans_word) {
  keyword == trans_word ||
    (nchar(keyword) >= 3 &&
     startsWith(trans_word, substr(keyword, 1, 3)) &&
     (startsWith(trans_word, keyword) || startsWith(keyword, trans_word)))
}
```


## 11.1 Phonological Distance of Transcription Errors

For accent-relevant keywords that participants got wrong, are their errors closer to the accented form or the standard form?

```{r metric1-phon-distance}
df_trans_main <- df_transcription_all %>% filter(batch_id == "Main Batch 3")

phon_dist_results <- list()

for (i in seq_len(nrow(df_trans_main))) {
  kws <- unlist(strsplit(tolower(gsub("[[:punct:]]", "", df_trans_main$Keywords[i])), "\\s+"))
  tws <- unlist(strsplit(tolower(gsub("[[:punct:]]", "", df_trans_main$Transcription[i])), "\\s+"))

  for (kw in kws) {
    if (!(kw %in% changed_keywords$keyword)) next

    ck <- changed_keywords[changed_keywords$keyword == kw, ]

    # Check if correctly matched (pmatch-style)
    matched <- any(vapply(tws, function(tw) is_pmatch(kw, tw), logical(1)))
    if (matched || length(tws) == 0) next

    # Keyword NOT matched — find closest word to the standard form
    dists_std <- as.integer(adist(tws, ck$standard))
    best_idx  <- which.min(dists_std)
    best_word <- tws[best_idx]

    d_std <- as.integer(adist(best_word, ck$standard))
    d_acc <- as.integer(adist(best_word, ck$accented))

    direction <- if (d_acc < d_std) "accent" else if (d_std < d_acc) "standard" else "equal"

    phon_dist_results[[length(phon_dist_results) + 1]] <- tibble(
      keyword = kw, contrast = ck$contrast,
      standard = ck$standard, accented = ck$accented,
      written = best_word,
      dist_to_standard = d_std, dist_to_accented = d_acc,
      direction = direction
    )
  }
}

phon_dist_df <- bind_rows(phon_dist_results)

cat(sprintf("Accent-relevant keywords that CHANGE under accent: %d\n", nrow(changed_keywords)))
cat(sprintf("Total incorrect transcriptions of these keywords: %d\n\n", nrow(phon_dist_df)))

# Direction summary
dir_counts <- table(phon_dist_df$direction)
total_err  <- nrow(phon_dist_df)

cat("Direction of errors:\n")
for (d in c("accent", "standard", "equal")) {
  n_d <- if (d %in% names(dir_counts)) dir_counts[[d]] else 0L
  cat(sprintf("  Closer to %-10s form: %d (%.1f%%)\n",
              toupper(d), n_d, n_d / max(total_err, 1) * 100))
}

# By contrast
cat("\nError direction by contrast:\n")
phon_dist_df %>%
  count(contrast, direction) %>%
  pivot_wider(names_from = direction, values_from = n, values_fill = 0) %>%
  mutate(total = rowSums(across(where(is.numeric)))) %>%
  arrange(match(contrast, contrast_order)) %>%
  knitr::kable()
```

```{r metric1-stats}
# Binomial test: proportion toward accent vs standard (excluding ties)
directional  <- phon_dist_df %>% filter(direction != "equal")
n_directional <- nrow(directional)
n_accent      <- sum(directional$direction == "accent")

if (n_directional > 0) {
  binom_result <- binom.test(n_accent, n_directional, p = 0.5)
  cat("Binomial test (accent vs standard, excluding ties):\n")
  cat(sprintf("  N = %d, toward accent = %d (%.1f%%)\n",
              n_directional, n_accent, n_accent / n_directional * 100))
  cat(sprintf("  p = %.4f\n\n", binom_result$p.value))
}

# Paired t-test: mean edit distance to standard vs accented
if (nrow(phon_dist_df) > 0) {
  t_result <- t.test(phon_dist_df$dist_to_standard,
                     phon_dist_df$dist_to_accented, paired = TRUE)
  cat(sprintf("Mean edit distance to standard: %.2f (SD = %.2f)\n",
              mean(phon_dist_df$dist_to_standard), sd(phon_dist_df$dist_to_standard)))
  cat(sprintf("Mean edit distance to accented:  %.2f (SD = %.2f)\n",
              mean(phon_dist_df$dist_to_accented), sd(phon_dist_df$dist_to_accented)))
  cat(sprintf("Paired t-test: t(%.0f) = %.3f, p = %.4f\n",
              t_result$parameter, t_result$statistic, t_result$p.value))
}

# Show example errors
cat("\nExample errors (first 15):\n")
phon_dist_df %>%
  head(15) %>%
  mutate(label = sprintf("  '%s' -> wrote '%s' (d_std=%d, d_acc=%d, -> %s)",
                          keyword, written, dist_to_standard, dist_to_accented, direction)) %>%
  pull(label) %>%
  cat(sep = "\n")
```


## 11.2 Accent-Distance Effect

```{r metric2-detection-by-contrast}
# A) Error detection rates by phonological contrast

has_phon_details <- "phon_err_details" %in% names(df_all)
detection_rates  <- NULL

if (has_phon_details) {
  trial_contrasts <- df_all %>%
    filter(stim_type == "target",
           !is.na(phon_err_details),
           phon_err_details != "", phon_err_details != "NA") %>%
    select(participant_id, question, phon_err_details) %>%
    distinct(participant_id, question, .keep_all = TRUE) %>%
    mutate(contrast = case_when(
      str_detect(phon_err_details, fixed("/ɪ/")) ~ "/ɪ/→/i/",
      str_detect(phon_err_details, fixed("/ʃ/")) ~ "/ʃ/→/s/",
      str_detect(phon_err_details, fixed("/k/")) ~ "/k/→/g/",
      str_detect(phon_err_details, fixed("/v/")) ~ "/v/→/b/",
      str_detect(phon_err_details, fixed("/ʊ/")) ~ "/ʊ/→/u/",
      str_detect(phon_err_details, fixed("/θ/")) ~ "/θ/→/t/",
      str_detect(phon_err_details, fixed("/i/")) ~ "/ɪ/→/i/",
      str_detect(phon_err_details, fixed("/g/")) ~ "/k/→/g/",
      str_detect(phon_err_details, fixed("/b/")) ~ "/v/→/b/",
      str_detect(phon_err_details, fixed("/u/")) ~ "/ʊ/→/u/",
      str_detect(phon_err_details, fixed("/t/")) ~ "/θ/→/t/",
      str_detect(phon_err_details, fixed("/s/")) ~ "/ʃ/→/s/",
      TRUE ~ NA_character_
    ))

  det_by_contrast <- df_trials %>%
    filter(condition %in% c("phon_err", "both_err")) %>%
    left_join(trial_contrasts, by = c("participant_id", "question")) %>%
    filter(!is.na(contrast))

  detection_rates <- det_by_contrast %>%
    group_by(contrast) %>%
    summarise(
      detected = sum(phon_detected_first, na.rm = TRUE),
      total    = n(),
      rate     = detected / total,
      .groups  = "drop"
    ) %>%
    mutate(contrast = factor(contrast, levels = contrast_order))

  cat("A) Error detection rates by phonological contrast\n\n")
  detection_rates %>% arrange(contrast) %>% knitr::kable(digits = 3)
} else {
  cat("Note: phon_err_details column not found — skipping detection-by-contrast.\n")
}
```

```{r metric2-transcription-by-contrast}
# B) Transcription accuracy of accent-relevant keywords by contrast

kw_by_contrast <- kw_df %>%
  filter(batch_id == "Main Batch 3") %>%
  inner_join(keyword_contrast, by = "keyword") %>%
  group_by(contrast) %>%
  summarise(
    correct  = sum(is_correct),
    total    = n(),
    accuracy = correct / total,
    .groups  = "drop"
  ) %>%
  mutate(contrast = factor(contrast, levels = contrast_order))

cat("\nB) Transcription accuracy of accent-relevant keywords by contrast\n\n")
kw_by_contrast %>% arrange(contrast) %>% knitr::kable(digits = 3)
```

```{r metric2-plot, fig.width=14, fig.height=6}
p_det <- NULL

if (!is.null(detection_rates) && nrow(detection_rates) > 0) {
  p_det <- ggplot(detection_rates, aes(x = contrast, y = rate, fill = contrast)) +
    geom_col(alpha = 0.8, color = "white") +
    scale_fill_manual(values = contrast_colors) +
    scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
    geom_text(aes(label = sprintf("%.2f", rate)), vjust = -0.5, size = 3.5) +
    labs(title = "Phonological Error Detection by Contrast",
         x = NULL, y = "Detection Rate") +
    theme(legend.position = "none")
}

p_trans <- ggplot(kw_by_contrast, aes(x = contrast, y = accuracy, fill = contrast)) +
  geom_col(alpha = 0.8, color = "white") +
  scale_fill_manual(values = contrast_colors) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  geom_text(aes(label = sprintf("%.2f", accuracy)), vjust = -0.5, size = 3.5) +
  labs(title = "Transcription Accuracy by Contrast",
       x = NULL, y = "Keyword Accuracy") +
  theme(legend.position = "none")

if (!is.null(p_det)) {
  p_det + p_trans + plot_layout(ncol = 2)
} else {
  p_trans
}
```

```{r metric2-correlation}
# C) Correlation between detection and transcription rates across contrasts
if (!is.null(detection_rates) && nrow(detection_rates) >= 4) {
  merged_rates <- detection_rates %>%
    select(contrast, det_rate = rate) %>%
    inner_join(kw_by_contrast %>% select(contrast, trans_acc = accuracy),
               by = "contrast")

  if (nrow(merged_rates) >= 4) {
    cor_result <- cor.test(merged_rates$det_rate, merged_rates$trans_acc)
    cat(sprintf("\nCorrelation: detection rate x transcription accuracy across contrasts\n"))
    cat(sprintf("  Pearson r = %.3f, p = %.4f (N = %d contrasts)\n",
                cor_result$estimate, cor_result$p.value, nrow(merged_rates)))
  }
}
```


## 11.3 Semantic Similarity

```{r metric3-semantic}
# Reconstruct intended sentence from filename; compute three similarity metrics

df_sim <- df_trans_main %>%
  mutate(
    intended       = str_replace(Filename, "\\.wav$", "") %>% str_replace_all("_", " "),
    trans_clean    = tolower(gsub("[[:punct:]]", "", Transcription)),
    intended_clean = tolower(gsub("[[:punct:]]", "", intended))
  ) %>%
  filter(trans_clean != "", !is.na(trans_clean))

cat(sprintf("Total trials with transcriptions: %d\n\n", nrow(df_sim)))

# ---- A) Jaccard word overlap ----
df_sim$jaccard <- mapply(function(int, trans) {
  i_words <- unique(unlist(strsplit(int, "\\s+")))
  t_words <- unique(unlist(strsplit(trans, "\\s+")))
  n_int   <- length(intersect(i_words, t_words))
  n_uni   <- length(union(i_words, t_words))
  if (n_uni > 0) n_int / n_uni else 0
}, df_sim$intended_clean, df_sim$trans_clean)

cat("A) Word Overlap (Jaccard Similarity)\n")
cat(sprintf("   M = %.3f, SD = %.3f\n", mean(df_sim$jaccard), sd(df_sim$jaccard)))
cat(sprintf("   Median = %.3f, Range = [%.3f, %.3f]\n\n",
            median(df_sim$jaccard), min(df_sim$jaccard), max(df_sim$jaccard)))

# ---- B) TF-IDF Cosine Similarity ----
n_trials <- nrow(df_sim)

docs <- tibble(
  doc_id = c(paste0("I_", seq_len(n_trials)), paste0("T_", seq_len(n_trials))),
  text   = c(df_sim$intended_clean, df_sim$trans_clean)
)

# Tokenise, count, compute TF-IDF via tidytext
doc_words <- docs %>%
  unnest_tokens(word, text) %>%
  count(doc_id, word) %>%
  bind_tf_idf(word, doc_id, n)

# Build TF-IDF matrix (docs x vocabulary)
doc_ids_ordered <- c(paste0("I_", seq_len(n_trials)), paste0("T_", seq_len(n_trials)))

tfidf_wide <- doc_words %>%
  select(doc_id, word, tf_idf) %>%
  pivot_wider(names_from = word, values_from = tf_idf, values_fill = 0)

tfidf_wide <- tfidf_wide[match(doc_ids_ordered, tfidf_wide$doc_id), ]
tfidf_mat  <- as.matrix(tfidf_wide[, -1])

# Cosine similarity for each intended–transcription pair
cosine_scores <- vapply(seq_len(n_trials), function(i) {
  v1    <- tfidf_mat[i, ]
  v2    <- tfidf_mat[n_trials + i, ]
  denom <- sqrt(sum(v1^2)) * sqrt(sum(v2^2))
  if (denom > 0) sum(v1 * v2) / denom else 0
}, numeric(1))

df_sim$cosine <- cosine_scores

cat("B) TF-IDF Cosine Similarity\n")
cat(sprintf("   M = %.3f, SD = %.3f\n", mean(cosine_scores), sd(cosine_scores)))
cat(sprintf("   Median = %.3f, Range = [%.3f, %.3f]\n\n",
            median(cosine_scores), min(cosine_scores), max(cosine_scores)))

# ---- C) Normalised Levenshtein Similarity ----
df_sim$lev_sim <- mapply(function(int, trans) {
  max_len <- max(nchar(int), nchar(trans))
  if (max_len == 0) return(1)
  1 - as.integer(adist(int, trans)) / max_len
}, df_sim$intended_clean, df_sim$trans_clean)

cat("C) Normalised Levenshtein Similarity\n")
cat(sprintf("   M = %.3f, SD = %.3f\n", mean(df_sim$lev_sim), sd(df_sim$lev_sim)))
cat(sprintf("   Median = %.3f, Range = [%.3f, %.3f]\n",
            median(df_sim$lev_sim), min(df_sim$lev_sim), max(df_sim$lev_sim)))
```

```{r metric3-participants-sentences}
# ---- D) Per-participant averages ----
pid_sim <- df_sim %>%
  group_by(participant_id) %>%
  summarise(
    jaccard_m = mean(jaccard), cosine_m = mean(cosine), lev_m = mean(lev_sim),
    .groups = "drop"
  )

cat(sprintf("\nD) Per-participant averages (N = %d)\n", nrow(pid_sim)))
cat(sprintf("   Jaccard: M = %.3f, SD = %.3f\n", mean(pid_sim$jaccard_m), sd(pid_sim$jaccard_m)))
cat(sprintf("   Cosine:  M = %.3f, SD = %.3f\n", mean(pid_sim$cosine_m),  sd(pid_sim$cosine_m)))
cat(sprintf("   Lev:     M = %.3f, SD = %.3f\n", mean(pid_sim$lev_m),     sd(pid_sim$lev_m)))

# ---- E) Per-sentence difficulty ----
# Use Filename as sentence identifier (strip .wav extension for labels)
df_sim <- df_sim %>%
  mutate(sentence_label = str_replace(Filename, "\\.wav$", ""))

sentence_sim <- df_sim %>%
  group_by(sentence_label) %>%
  summarise(cosine_m = mean(cosine), jaccard_m = mean(jaccard), n = n(),
            .groups = "drop") %>%
  arrange(cosine_m)

cat("\nE) Per-sentence semantic similarity (by TF-IDF cosine)\n")
sentence_sim %>% knitr::kable(digits = 3)
```

```{r metric3-kw-correlation}
# ---- F) Keyword accuracy x Semantic similarity ----
cor_kw_cos    <- cor.test(df_sim$PropKeywordsCorrect, df_sim$cosine)
cor_kw_cos_sp <- cor.test(df_sim$PropKeywordsCorrect, df_sim$cosine, method = "spearman")

cat(sprintf("\nF) Keyword accuracy x Semantic similarity\n"))
cat(sprintf("   Pearson  r = %.3f, p = %.6f (N = %d trials)\n",
            cor_kw_cos$estimate, cor_kw_cos$p.value, nrow(df_sim)))
cat(sprintf("   Spearman rho = %.3f, p = %.6f\n",
            cor_kw_cos_sp$estimate, cor_kw_cos_sp$p.value))
```

```{r metric3-plot, fig.width=18, fig.height=5}
p_hist <- ggplot(df_sim, aes(x = cosine)) +
  geom_histogram(bins = 25, fill = COL_PHON, alpha = 0.7, color = "white") +
  geom_vline(xintercept = mean(df_sim$cosine), linetype = "dashed", linewidth = 1) +
  annotate("text", x = mean(df_sim$cosine) + 0.02, y = Inf, vjust = 2,
           label = sprintf("M = %.3f", mean(df_sim$cosine)), size = 4) +
  labs(title = "Distribution of Semantic Similarity",
       x = "TF-IDF Cosine Similarity", y = "Number of Trials")

p_scatter <- ggplot(df_sim, aes(x = PropKeywordsCorrect, y = cosine)) +
  geom_point(alpha = 0.3, color = COL_NONPHON, size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "grey40", fill = "grey80") +
  labs(title = sprintf("Keyword Accuracy x Semantic Similarity (r = %.3f)",
                        cor_kw_cos$estimate),
       x = "Keyword Accuracy (PropCorrect)",
       y = "TF-IDF Cosine Similarity")

p_sentence <- NULL
if (nrow(sentence_sim) > 0) {
  p_sentence <- ggplot(sentence_sim,
                        aes(x = cosine_m, y = reorder(sentence_label, cosine_m))) +
    geom_col(fill = COL_CORRECT, alpha = 0.7, color = "white") +
    labs(title = "Semantic Similarity by Sentence",
         x = "Mean Cosine Similarity", y = NULL) +
    scale_x_continuous(limits = c(0, 1))
}

if (!is.null(p_sentence)) {
  p_hist + p_scatter + p_sentence + plot_layout(ncol = 3)
} else {
  p_hist + p_scatter + plot_layout(ncol = 2)
}
```


# Session Info

```{r session-info}
sessionInfo()
```
