---
title: "Data Preprocessing"
subtitle: "Exp 1: Error Detection in Spanish-Accented English"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.width = 12,
                      fig.height = 6,
                      dev = "png",
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE,
                      cache = FALSE,
                      include = TRUE)

options(repos = c(CRAN = "https://cloud.r-project.org"))
```

```{r libraries, echo=FALSE, include=FALSE}
rm(list = ls())

library(tidyverse)
library(magrittr)
library(stringr)
library(forcats)
library(fs)        # for dir_ls() in demographics section
```

```{r constants, echo=FALSE, include=FALSE}
dodge_amt <- 0.9
dodge_amt2 <- 0.5

theme_set(
  theme_minimal(base_size = 20) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20),
        axis.ticks = element_line(color = "black"),
        axis.line = element_blank(),
        axis.line.x.bottom = element_line(color = "black"),
        axis.line.y.left = element_line(color = "black"),
        strip.background = element_rect(fill = "white", color = "black")))
```


# Core preprocessing function

This function takes a raw Gorilla CSV export and a stimulus design file, filters to eligible participants who passed the sanity check, extracts trial metadata (item_base, item_type, condition), joins stimulus info, and prints diagnostic summaries. Returns a list of data frames split by block (all, error_detection, questionnaire, transcription).

```{r function-process-experiment-data}
process_experiment_data <- function(data_filepath, stimulus_filepath) {

  data <- read_csv(data_filepath)

  # Initial inspection
  print(colnames(data))
  print(paste("N =", n_distinct(data$participant_id)))
  print(unique(data$block_name))
  print(unique(data$branch_eligiblenotEligible))

  # Filter to eligible participants.
  # The sanity check column (branch_sanitycheckFailsanitycheckPass) was only
  # added starting in batch 3. For earlier batches, we skip that filter.
  data_clean <- data %>%
    filter(branch_eligiblenotEligible == "eligible")

  if ("branch_sanitycheckFailsanitycheckPass" %in% colnames(data_clean)) {
    data_clean <- data_clean %>%
      filter(branch_sanitycheckFailsanitycheckPass == "sanitycheckPass")
    print("Sanity check filter applied.")
  } else {
    print("No sanity check column found (pre-batch 3 data); skipping that filter.")
  }

  print(paste("N of Eligible Participants =", n_distinct(data_clean$participant_id)))

  # Define attention check items
  att_check_good_items <- c("fill_25", "fill_26", "fill_27", "fill_28")
  att_check_err_items <- c("fill_29", "fill_30", "fill_31", "fill_32")

  # Extract trial metadata from stimuli_presented strings
  d_processed <- data_clean %>%
    mutate(
      Block = case_when(
        str_starts(block_name, "Trivia_List") ~ "Error_Detection",
        block_name == "Questionnaire" ~ "Questionnaire",
        block_name == "Transcription" ~ "Transcription",
        TRUE ~ block_name
      ),
      # Extract item identifier (e.g., targ_01, fill_05)
      item_base = str_extract(stimuli_presented, "targ_\\d+|fill_\\d+"),
      # Classify item type
      item_type = case_when(
        str_starts(item_base, "targ_") ~ "target",
        item_base %in% att_check_good_items ~ "att_check_good",
        item_base %in% att_check_err_items ~ "att_check_err",
        str_starts(item_base, "fill_") ~ "filler_other",
        TRUE ~ "other_non_err_detection"
      ),
      # Extract condition letter (A/B/C/D) from filename
      cond_letter = str_extract(stimuli_presented, "(?<=_)[ABCD](?=_)"),
      Condition = case_when(
        cond_letter == "A" ~ "correct",
        cond_letter == "B" ~ "phon_err",
        cond_letter == "C" ~ "non_phon_err",
        cond_letter == "D" ~ "both_err",
        TRUE ~ NA_character_
      ),
      trial_duration_num = as.numeric(trial_duration),
      response_rt_num = as.numeric(response_rt)
    )

  # Join with stimulus design file to get error details
  stim_design <- read_csv(stimulus_filepath)

  stim_info <- stim_design %>%
    mutate(PhonErr_Details = str_c(phon_err_type, word_pairs, sep = "; ")) %>%
    select(ID, nonphon_err_type, PhonErr_Details, Answer)

  d_processed <- d_processed %>%
    left_join(stim_info, by = c("item_base" = "ID"))

  # Diagnostic: verify join worked for targets
  print(
    d_processed %>%
      select(participant_id, item_base, Condition, nonphon_err_type, PhonErr_Details) %>%
      filter(str_starts(item_base, "targ_")) %>%
      head()
  )

  # Participants per trivia list (should be balanced across 4 lists)
  participants_per_list <- d_processed %>%
    filter(Block == "Error_Detection") %>%
    distinct(participant_id, block_name) %>%
    group_by(block_name) %>%
    summarise(n_participants = n(), .groups = 'drop')
  print(participants_per_list)

  # Trials per condition per list (should be 6 targets per condition)
  unique_trial_tally <- d_processed %>%
    filter(Block == "Error_Detection") %>%
    filter(item_type != "other_non_trivia") %>%
    distinct(block_name, participant_id, item_base, item_type, Condition) %>%
    group_by(block_name, item_type, Condition) %>%
    summarise(n_unique_trials_total = n(), .groups = 'drop')

  list_summary <- unique_trial_tally %>%
    left_join(participants_per_list, by = "block_name") %>%
    mutate(trials_per_participant = n_unique_trials_total / n_participants) %>%
    select(block_name, item_type, Condition, trials_per_participant,
           n_unique_trials_total, n_participants) %>%
    arrange(block_name, item_type, Condition)
  print(list_summary, n = Inf)

  # Transcription trial summary
  participants_transcription <- d_processed %>%
    filter(Block == "Transcription") %>%
    distinct(participant_id, block_name) %>%
    group_by(block_name) %>%
    summarise(n_participants = n(), .groups = 'drop')

  unique_trials_transcription <- d_processed %>%
    filter(Block == "Transcription") %>%
    filter(!is.na(trial_num)) %>%
    distinct(participant_id, block_name, trial_num, trial_template)

  transcription_total_tally <- unique_trials_transcription %>%
    group_by(block_name, trial_template) %>%
    summarise(n_unique_trials_total = n(), .groups = 'drop')

  transcription_summary <- transcription_total_tally %>%
    left_join(participants_transcription, by = "block_name") %>%
    mutate(trials_per_participant = n_unique_trials_total / n_participants) %>%
    select(block_name, trial_template, trials_per_participant,
           n_unique_trials_total, n_participants) %>%
    arrange(block_name, trial_template)
  print(transcription_summary, n = Inf)

  # Select and rename columns for final output; filter to task blocks only
  rows_to_exclude <- c(
    "Transcription_Instr",
    "volumeAdjusted|playTranscription|volumeAdjustPressC",
    "End_Main_Experiment",
    "questionnaire_part1_begin",
    "questionnaire_part2_begin",
    "confirmationCode",
    "End_Study"
  )

  d.final <- d_processed %>%
    filter(Block %in% c("Error_Detection", "Transcription", "Questionnaire")) %>%
    select(
      group_id, participant_id, response_value, stimuli_presented,
      stimulus_files_presented, trial_num, Block, item_base, item_type,
      Condition, nonphon_err_type, PhonErr_Details, Answer
    ) %>%
    rename(
      block = Block,
      condition = Condition,
      phon_err_details = PhonErr_Details,
      answer = Answer
    ) %>%
    filter(!stimuli_presented %in% rows_to_exclude) %>%
    mutate(
      # Code whether participant chose to re-listen on second-attempt screens
      second_attempt_used = case_when(
        str_detect(stimuli_presented, "Second_attempt") &
          response_value != "I'm good — skip to the next question" ~ "Yes",
        str_detect(stimuli_presented, "Second_attempt") &
          response_value == "I'm good — skip to the next question" ~ "No",
        TRUE ~ "No"
      )
    ) %>%
    group_by(participant_id, block) %>%
    arrange(trial_num, .by_group = TRUE) %>%
    mutate(
      # Create sequential trial numbers within each block
      trial_num_sequential = case_when(
        str_starts(block, "Error_Detection") ~ dense_rank(item_base),
        TRUE ~ row_number()
      )
    ) %>%
    ungroup() %>%
    select(-trial_num)

  # Final sanity checks
  participant_count_final <- n_distinct(d.final$participant_id)
  print(paste("Total unique participants in final data:", participant_count_final))

  trial_counts <- d.final %>%
    group_by(participant_id, block) %>%
    summarise(n_trials = n_distinct(trial_num_sequential), .groups = 'drop')
  print("Trial counts per participant per block:")
  print(trial_counts)

  block_counts <- d.final %>%
    distinct(participant_id, block) %>%
    group_by(block) %>%
    summarise(n_participants = n(), .groups = 'drop')
  print("Participants per block:")
  print(block_counts)

  # Add numeric participant ID for easy reference
  d.final <- d.final %>%
    mutate(participant_num = as.integer(as.factor(participant_id)))

  # Split into separate data frames by block
  d_err_detection <- d.final %>% filter(str_detect(block, "Error_Detection"))
  d_questionnaire <- d.final %>% filter(str_detect(block, "Questionnaire"))
  d_transcription <- d.final %>% filter(str_detect(block, "Transcription"))

  return(list(
    all_data = d.final,
    err_detection_data = d_err_detection,
    questionnaire_data = d_questionnaire,
    transcription_data = d_transcription
  ))
}
```

# Helper: reformat batch 1 first-pass data

Batch 1 had a randomization issue where first and second attempts were interleaved in the raw data. This function identifies true first-pass trials and sorts them with attention checks on top (for RA coding).

```{r function-reformat-first-pass}
reformat_first_pass <- function(err_detection_df) {

  d.batch1.original <- err_detection_df

  # Mark true first pass: every 2 rows per item = 1 attempt
  d.marked <- d.batch1.original %>%
    group_by(participant_id, item_base) %>%
    arrange(trial_num_sequential) %>%
    mutate(true_attempt_num = (row_number() - 1) %/% 2 + 1) %>%
    ungroup()

  # Sanity checks
  check_data <- d.marked %>%
    filter(grepl("Second_attempt", stimuli_presented)) %>%
    select(participant_id, item_base, trial_num_sequential, stimuli_presented, true_attempt_num)
  print(head(check_data))

  target_check_data <- d.marked %>%
    filter(!grepl("fill_", item_base)) %>%
    select(participant_id, item_base, trial_num_sequential, stimuli_presented, true_attempt_num)
  print(head(target_check_data))

  # Keep only first-pass trials
  d.first_pass_only <- d.marked %>%
    filter(true_attempt_num == 1)

  # Sort: attention checks first, then targets, then fillers
  d.first_pass_sorted <- d.first_pass_only %>%
    mutate(
      sort_category = case_when(
        startsWith(item_type, "att_check_") ~ 1,
        !grepl("fill_", item_base)         ~ 2,
        TRUE                               ~ 3
      )
    ) %>%
    arrange(participant_id, sort_category, trial_num_sequential)

  # Reassign sequential participant numbers
  participant_map <- d.first_pass_sorted %>%
    distinct(participant_id) %>%
    arrange(participant_id) %>%
    mutate(participant_num = row_number())

  d.final_with_num <- d.first_pass_sorted %>%
    left_join(participant_map, by = "participant_id") %>%
    select(participant_num, participant_id, everything())

  return(list(
    all_attempts_marked = d.marked,
    first_pass_only_unsorted = d.first_pass_only,
    first_pass_only_sorted = d.final_with_num
  ))
}
```

# Helper: save preprocessed files

```{r function-save-preprocessed}
save_preprocessed <- function(output_dir, prefix,
                               all_data = NULL,
                               err_data = NULL,
                               quest_data = NULL,
                               transcrip_data = NULL,
                               save_attcheck_top = FALSE) {

  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
    print(paste("Created directory:", output_dir))
  }

  if (!is.null(all_data)) {
    write_csv(all_data, file.path(output_dir, paste0(prefix, "_all_data.csv")))
  }
  if (!is.null(err_data)) {
    write_csv(err_data, file.path(output_dir, paste0(prefix, "_err_detection.csv")))

    # Optionally save version with attention checks sorted to top (for RA coding)
    if (save_attcheck_top) {
      err_att_top <- err_data %>%
        group_by(participant_id) %>%
        arrange(desc(str_detect(item_type, "^att_check")), .by_group = TRUE) %>%
        ungroup()
      write_csv(err_att_top, file.path(output_dir, paste0(prefix, "_err_detection_attcheck_top.csv")))
    }
  }
  if (!is.null(quest_data)) {
    write_csv(quest_data, file.path(output_dir, paste0(prefix, "_questionnaire_data.csv")))
  }
  if (!is.null(transcrip_data)) {
    write_csv(transcrip_data, file.path(output_dir, paste0(prefix, "_transcription_data.csv")))
  }

  print(paste("Saved preprocessed files to:", output_dir))
}
```


# Batch 1: 48 participants (failed randomization)

These participants had a randomization bug where first and second attempts were not properly separated. We use `reformat_first_pass()` to extract only the first-pass data.

```{r batch1-48subj}
d_main_batch1_48subj <- process_experiment_data(
  data_filepath = "../data/raw/main_110425_random_fail_48subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

# Extract first-pass error detection data
first_pass_list <- reformat_first_pass(d_main_batch1_48subj$err_detection_data)
df_err_batch1_48subj_first_pass <- first_pass_list$first_pass_only_sorted

# Save all outputs
save_preprocessed(
  output_dir = "../data/preprocessed/main_batch1/",
  prefix = "48subj",
  all_data = d_main_batch1_48subj$all_data,
  err_data = d_main_batch1_48subj$err_detection_data,
  quest_data = d_main_batch1_48subj$questionnaire_data,
  transcrip_data = d_main_batch1_48subj$transcription_data
)

# Save first-pass-only error detection separately
write_csv(
  df_err_batch1_48subj_first_pass,
  "../data/preprocessed/main_batch1/48subj_err_detection_first_pass.csv"
)
```


# Batch 1: 4 participants (test run, fixed order)

```{r batch1-4subj-testrun}
d_main_batch1_4subj <- process_experiment_data(
  data_filepath = "../data/raw/main_102825_test_run_4_subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

save_preprocessed(
  output_dir = "../data/preprocessed/main_batch1/",
  prefix = "testrun_4subj",
  all_data = d_main_batch1_4subj$all_data,
  err_data = d_main_batch1_4subj$err_detection_data,
  quest_data = d_main_batch1_4subj$questionnaire_data,
  transcrip_data = d_main_batch1_4subj$transcription_data
)
```


# Batch 2: 4 participants (test run, fixed randomization)

```{r batch2-4subj-testrun}
d_main_batch2_4subj <- process_experiment_data(
  data_filepath = "../data/raw/main_110725_random_fixed_test_run_4subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

save_preprocessed(
  output_dir = "../data/preprocessed/main_batch2/",
  prefix = "testrun_4subj",
  all_data = d_main_batch2_4subj$all_data,
  err_data = d_main_batch2_4subj$err_detection_data,
  quest_data = d_main_batch2_4subj$questionnaire_data,
  transcrip_data = d_main_batch2_4subj$transcription_data
)
```


# Transcription baseline: 24 participants (batches 1+2 combined)

These participants did transcription only (no error detection task), serving as the baseline comparison group.

```{r transcription-baseline}
d_transcrip_baseline_batch1 <- process_experiment_data(
  data_filepath = "../data/raw/transcription_baseline_batch1_14subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

d_transcrip_baseline_batch2 <- process_experiment_data(
  data_filepath = "../data/raw/transcription_baseline_batch2_10subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

# Combine the two baseline batches
df_transcrip_baseline_all <- bind_rows(
  d_transcrip_baseline_batch1$all_data,
  d_transcrip_baseline_batch2$all_data
)
df_transcrip_transcrip_baseline_all <- bind_rows(
  d_transcrip_baseline_batch1$transcription_data,
  d_transcrip_baseline_batch2$transcription_data
)
df_quest_transcrip_baseline_all <- bind_rows(
  d_transcrip_baseline_batch1$questionnaire_data,
  d_transcrip_baseline_batch2$questionnaire_data
)

save_preprocessed(
  output_dir = "../data/preprocessed/transcrip_baseline/",
  prefix = "",
  all_data = df_transcrip_baseline_all,
  quest_data = df_quest_transcrip_baseline_all,
  transcrip_data = df_transcrip_transcrip_baseline_all
)
```


# Batch 3: 4 participants (test run, correct randomization)

```{r batch3-4subj-testrun}
d_main_batch3_4subj <- process_experiment_data(
  data_filepath = "../data/raw/main_112825_test_run_4subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

save_preprocessed(
  output_dir = "../data/preprocessed/main_batch3/",
  prefix = "testrun_4subj",
  all_data = d_main_batch3_4subj$all_data,
  err_data = d_main_batch3_4subj$err_detection_data,
  quest_data = d_main_batch3_4subj$questionnaire_data,
  transcrip_data = d_main_batch3_4subj$transcription_data,
  save_attcheck_top = TRUE
)
```


# Batch 3: 48 participants (correct randomization) -- PRIMARY ANALYSIS DATA

This is the main dataset used in all Exp 1 analyses. Correct Latin-Square randomization across 4 trivia lists, 6 targets per
condition per participant, 8 attention check fillers.

```{r batch3-48subj}
d_main_batch3_48subj <- process_experiment_data(
  data_filepath = "../data/raw/main_120325_48subj.csv",
  stimulus_filepath = "../data/Stimuli_design.csv"
)

save_preprocessed(
  output_dir = "../data/preprocessed/main_batch3/",
  prefix = "48subj",
  all_data = d_main_batch3_48subj$all_data,
  err_data = d_main_batch3_48subj$err_detection_data,
  quest_data = d_main_batch3_48subj$questionnaire_data,
  transcrip_data = d_main_batch3_48subj$transcription_data,
  save_attcheck_top = TRUE
)
```


# Demographics summary

Combines all demographic data from Prolific exports across batches.

```{r demographics-summary, echo=FALSE, warning=FALSE, message=FALSE}
file_list <- dir_ls(path = "../data/demographic", glob = "*.csv")

# Skip the combined file if it already exists (it's our own output)
file_list <- file_list[!str_detect(file_list, "combined_demographics")]

process_file <- function(file_path) {
  df <- read_csv(file_path, show_col_types = FALSE)

  # Raw Prolific exports have a Status column; already-processed files may not
  if ("Status" %in% colnames(df)) {
    df <- df %>% filter(Status == "APPROVED")
  }

  # Check which column names are available and select accordingly
  if ("Sex" %in% colnames(df)) {
    df <- df %>%
      select(
        gender = Sex,
        race = `Ethnicity simplified`,
        age = Age
      ) %>%
      mutate(ethnicity = race)
  } else if ("gender" %in% colnames(df)) {
    # Already in processed format
    df <- df %>% select(any_of(c("gender", "race", "age", "ethnicity")))
  }

  return(df)
}

df_all_demo <- map(file_list, process_file) %>% list_rbind()

write_csv(df_all_demo, "../data/demographic/combined_demographics.csv")

print(paste("Total participants in demographics:", nrow(df_all_demo)))
print("Gender breakdown:")
print(table(df_all_demo$gender))
print("Age summary:")
print(summary(df_all_demo$age))
```
